---
title: "final ntw results"
date: "2026-05-24"
---

# preamble and setup

```{r pathfinding, message=FALSE}
here::i_am("scripts/_mss/final-res.Rmd")
```

for collated viz + stats results for ntw mss

-   notes on the viz:
    -   main viz = focuses on trt type (NTW vs control), minimum temp, population
    -   supp viz = includes year and sex (if applicable)
-   notes on the stats:
    -   main stats = focuses on the NTW bugs and population, minimum temp as predictors
    
main results/figs:

1. survival: hatch to pupation; hatch to eclosion
2. mass: pupal mass; eclosion mass
3. development time: hatch to pupation, pupation to eclosion
4. adult longevity (2023 vs 2025)
5. adult fitness (2023 vs 2025)


# loading

```{r data, message=FALSE}
source(here::here("scripts/_mss/load.R"))
```

## plot objects
```{r}
# for saving/exporting patchworks
p <- list() # main plots
ps <- list() # supp plots
```

# ## development

# (1) survival

## survival to pup

```{r plot prep pup survival}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = prop.pup,
    ) +
    geom_line(show.legend = TRUE) +
    geom_point(size = 2,
               #aes(size = n)
               ) +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    #scale_radius(range = c(2,4)) +
    ylim(c(0, 1))
}
```

```{r main fig pup survival}
p <- list_modify(
  p,
  surv_pup =
    dfs_viz$wide %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps() %>% #ResToCsv("dev/ss/surv")
    .plotfn() +
    labs(#y = "% survived to pupation",
      title = "hatch to pupation"
    )
)

p$surv_pup
```

### supp
```{r supp fig pup survival}
ps <- list_modify(
  ps,
  surv_pup =
    dfs_viz$wide %>%
    CalcSurvProps_def() %>% #ResToCsv("dev/ss/surv-supp")
    .plotfn() +
    labs(y = "% survival to pupation") +
    facet_wrap(~year)
)

ps$surv_pup
```

### stats
```{r main stats pup survival}
mods <- list(
  m1 = glm(is.pup ~ year*pop*trt.minT,
           family = "binomial", data = dfs_stats$dev_expt),
  m1a = glm(is.pup ~ year*pop*trt.minT + as.factor(is.sup),
           family = "binomial", data = dfs_stats$dev_expt)
)

AIC(mods$m1, mods$m1a) # technically better w sup but also lots of missing obs
anova(mods$m1, test = "Chisq") # can drop the 3-way

mods <- list_modify(
  mods,
  m2 = glm(is.pup ~ (year + pop + trt.minT)^2,
           family = "binomial", data = dfs_stats$dev_expt)
)

anova(mods$m2, test = "Chisq") #%>% ResToCsv("dev/anova/surv-pup")
summary(mods$m2)
```

## survival to eclosion
```{r plot prep ec survival}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = prop.ec) +
    geom_line(show.legend = TRUE) +
    geom_point(size = 2,) +
    geom_errorbar(aes(ymax = prop.ec + se.ec,
                      ymin = prop.ec - se.ec),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    ylim(c(0, 1))
}
```


```{r main fig adult survival}
# TODO include sample size?
p <- list_modify(
  p,
  surv_ec = 
    dfs_viz$wide %>%
    filter(is.na(is.LPI),
           #!is.na(sex) # can keep?
           ) %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps() %>% 
    .plotfn() +
    labs(#y = "% pupa eclosed",
         title = "pupa to eclosion",
         )
)

p$surv_ec
```

### supp

```{r supp fig ec survival}
ps <- list_modify(
  ps,
  surv_ec = 
    dfs_viz$wide %>%
    CalcSurvProps_def() %>% 
    .plotfn() +
    #geom_point(aes(size = sqrt(n))) +
    labs(y = "% survival to eclosion") +
    facet_wrap(~year)
)

ps$surv_ec
```

### stats

```{r main stats surv to ec}
# filter df bc of including mass.pupa in the model -- m1
# will fit extra observations
df <- dfs_stats$dev_expt %>%
  filter(!is.na(mass.pupa))

mods <- list(
  m1 = glm(is.ec ~ year*pop*trt.minT,
           family = "binomial", data = df),
  m1a = glm(is.ec ~ year*pop*trt.minT + mass.pupa,
           family = "binomial", data = df)
)

AIC(mods$m1, mods$m1a) 
anova(mods$m1, mods$m1a, test = "Chisq")
# including mass is better, per anova + aic

#anova(mods$m1, test = "Chisq") # just year and pop matter lmao
anova(mods$m1a, test = "Chisq") # 3 way is marginally signif

# try to simplify?
mods <- list_modify(
  mods,
  m2 = glm(is.ec ~ (year + pop + trt.minT)^2 + mass.pupa,
           family = "binomial", data = df),
)

AIC(mods$m2, mods$m1a) 
anova(mods$m2, mods$m1a, test = "Chisq")
# more complex model is slightly better, per AIC

anova(mods$m1a, test = "Chisq") #%>% ResToCsv("dev/anova/surv-ec")
summary(mods$m1a)
```

```{r}
# looking at pup mass
dfs_viz$wide %>%
    filter(is.na(is.LPI),
           !is.na(is.ec)
           ) %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
  group_by(is.ec, .add = TRUE) %>%
  summarise(n = n(),
            avg.wmass = mean(mass.wander, na.rm = TRUE),
            se.wmass= se(mass.wander),
            avg.pmass = mean(mass.pupa, na.rm = TRUE),
            se.pmass = se(mass.pupa)) %>% View()
```


## collated figs

```{r}
# main
p$surv_pup + theme(legend.position = "none") + 
  p$surv_ec +
  plot_layout(axes = "collect", guides = "collect") &
  labs(y = "% survival") &
  theme(plot.title = element_text(hjust = 0.5))  

#ResToFig("dev-surv")

# supp
(ps$surv_pup + 
    theme(legend.position = "top",
                     legend.margin = margin()) +
    labs(y = "% survival\nto pupation",
                    x = element_blank())
  ) / 
  (ps$surv_ec + 
     labs(y = "% survival\nto eclosion") +
     guides(col = "none", shape = "none", lty = "none") +
     theme(legend.position = "none")
   ) +
  plot_layout(axes = "collect", #guides = "collect"
              )

#ResToFig("supp/dev-surv")
```

## sample sizes
```{r}
dfs_viz$wide %>%
  group_by(year, pop, trt) %>%
  summarise(n = sum(!is.na(is.pup)),
            wand = sum(!is.na(jdate.wander), na.rm = TRUE),
            pup = sum(is.pup == 1, na.rm = TRUE),
            ec = sum(is.ec == 1, na.rm = TRUE),
            LPI = sum(is.pup == 2, na.rm = TRUE)) #%>% ResToCsv("dev/N-surv")
```


# (2) mass 

```{r plot fn mass}
.plotfn <- function(x) {
  x %>%
    Plot_Dev() +
    aes(y = avg.mass) +
    geom_point(size = 2) +
    geom_line(show.legend = TRUE) +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass),
                  width = 0.5, lty = "solid", show.legend = FALSE,
                  ) +
    labs(y = "mass (g)")
}
```

## pupal mass

drop LPIs bc mass isnt accurate

```{r df prep pup mass}
df <- dfs_viz$long %>%
  filter(!is.na(sex),
         instar == "pupa")
```

```{r main figs pup mass}
p <- list_modify(
  p,
  mass_pup = # sexes combined
    df %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    .plotfn() +
    labs(title = "pupa"),
  
  mass_pupsex = # sexes parsed
    df %>%
    PrepAdultSS() %>%
    group_by(sex, .add = TRUE) %>%
    ungroup(year) %>%
    CalcDevSS() %>% #ResToCsv("dev/ss/mass-pup-sex")
    .plotfn() +
    facet_wrap(~ sex, labeller = as_labeller(p_scales$facs_sex)) +
    labs(title = "pupa") 
)

p$mass_pup
p$mass_pupsex # use this, per stats
```

### supp

```{r supp figs pup mass}
ps <- list_modify(
  ps,
  # sexes combined
  mass_pup =
    df %>%
    PrepAdultSS() %>%
    ungroup(starts_with("trt")) %>% 
    group_by(trt, trt.type, trt.minT, .add = TRUE) %>%
    CalcDevSS() %>% #View()
    .plotfn() +
    facet_wrap(~ year) +
    ylab("pupal mass (g)"),
  
  # sexes parsed (use this one)
  mass_pupsex =
  df %>%
    PrepAdultSS() %>%
    ungroup(starts_with("trt")) %>% 
    group_by(trt, trt.type, trt.minT, sex, .add = TRUE) %>% 
    CalcDevSS() %>% #ResToCsv("dev/ss/mass-pup-sex-supp") 
    #filter(trt != 260) %>%
    .plotfn() +
    facet_grid(sex ~ year, labeller = labeller(sex = c(p_scales$facs_sex)),
               #scales = "free"
               ) +
    ylab("pupal mass (g)")
)

ps$mass_pup
ps$mass_pupsex
```


### stats
```{r main stats pup mass}
mods <- list(
  m1 = lm(mass.pupa ~ year*pop*trt.minT*sex,
           data = dfs_stats$dev_expt
          )
)

anova(mods$m1) # ok to just focus on the 2 ways of trt/pop/year and main effect of sex

mods <- list_modify(
  mods,
  m2 = lm(mass.pupa ~ (year + pop + trt.minT + sex)^2,
           data = dfs_stats$dev_expt),
  m3 = lm(mass.pupa ~ (year + pop + trt.minT)^2 + sex,
           data = dfs_stats$dev_expt)
)

anova(mods$m3, mods$m2, mods$m1)
## adding model complexity doesnt do much
AIC(mods$m3, mods$m2, mods$m1)
## m3 = lowest AIC

anova(mods$m3) #%>% ResToCsv("dev/anova/mass-pup")
summary(mods$m3)
```

## eclosion mass

```{r df prep eclosion mass}
df <- dfs_viz$long %>%
  filter(!is.na(sex),
         instar == "eclose")
```

```{r main fig eclosion mass}
p <- list_modify(
  p,
  mass_ec = # sexes combined
    df %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    .plotfn() +
    labs(title = "adult"),
  
  mass_ecsex = # sexes parsed
    df %>%
    PrepAdultSS() %>%
    group_by(sex, .add = TRUE) %>%
    ungroup(year) %>%
    CalcDevSS() %>% #ResToCsv("dev/ss/mass-ec-sex")
    .plotfn() +
    facet_wrap(~ sex, labeller = as_labeller(p_scales$facs_sex)) +
    labs(title = "adult")
)

p$mass_ec
p$mass_ecsex
```

### supp

```{r supp figs eclosion mass}
ps <- list_modify(
  ps,
  # sexes combined
  mass_ec =
    df %>%
    PrepAdultSS() %>%
    ungroup(starts_with("trt")) %>% 
    group_by(trt, trt.type, trt.minT, .add = TRUE) %>%
    CalcDevSS() %>%
    .plotfn() +
    facet_wrap(~ year) +
    ylab("adult mass (g)"),
  
  # sexes parsed 
  mass_ecsex =
    df %>%
    PrepAdultSS() %>%
    ungroup(starts_with("trt")) %>% 
    group_by(trt, trt.type, trt.minT, sex, .add = TRUE) %>%
    CalcDevSS() %>% #ResToCsv("dev/ss/mass-ec-sex-supp")
    #filter(trt != 260) %>%
    .plotfn() +
    facet_grid(sex ~ year, labeller = labeller(sex = c(p_scales$facs_sex)),
               #scales = "free"
               ) +
    ylab("adult mass (g)")
)

ps$mass_ec
ps$mass_ecsex
```

### stats

```{r}
# include pupal mass?
mods <- list(
    m1 = lm(mass.eclose ~ year*pop*trt.minT*sex + mass.pupa,
           data = dfs_stats$dev_expt),
    m1a = lm(mass.eclose ~ year*pop*trt.minT*sex,
            data = dfs_stats$dev_expt),
    m1b = lm(mass.eclose ~ year*pop*trt.minT + mass.pupa,
            data = dfs_stats$dev_expt)
)

AIC(mods$m1, mods$m1a, mods$m1b) # yes, based on AIC

# which interactions to keep?
anova(mods$m1) # main effs, and 2-ways with minT:sex, year:sex

mods <- list_modify(
  mods,
  m3 = lm(mass.eclose ~ (year + pop + trt.minT + sex)^2 + mass.pupa,
          data = dfs_stats$dev_expt),
  m4 = lm(mass.eclose ~ (year + pop + trt.minT)^2 + sex + mass.pupa,
          data = dfs_stats$dev_expt),
  m5 = lm(mass.eclose ~ year*sex + pop + trt.minT*sex + mass.pupa,
          data = dfs_stats$dev_expt)
)

anova(mods$m3)
anova(mods$m4)
AIC(mods$m1, mods$m3, mods$m4, mods$m5) # isolating the 2ways is better

anova(mods$m3, mods$m4) #m4 better
anova(mods$m4, mods$m5) #m5 not better than m4?
anova(mods$m3, mods$m4, mods$m5)

# just keep the one with the 2 ways bc way easier to explain in methods lol
anova(mods$m3) #%>% ResToCsv("dev/anova/mass-ec")
summary(mods$m3)
```



## collated figs

```{r}

# main
p$mass_pupsex + theme(legend.position = "none") +
  p$mass_ecsex +
  plot_layout(axes = "collect", #guides = "collect"
              widths = c(0.9, 1)) &
  labs(x = p_scales$xlab_minT) &
  theme(plot.title = element_text(hjust = 0.5))

#ResToFig("dev-mass")

# supp
(ps$mass_pupsex +
    theme(legend.position = "top",
          legend.margin = margin()) +
    labs(x = element_blank())) / 
  (ps$mass_ecsex +
     theme(legend.position = "none") +
     guides(col = "none", lty = "none", shape = "none"))  +
  plot_layout(axes = "collect", #guides = "collect"
              )

#ResToFig("supp/dev-mass")
```

# (3) development time

```{r plot prep dev time}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = avg.tt) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.tt + se.tt,
                      ymin = avg.tt - se.tt),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "duration (days)")
}
```

## time to pupation
```{r main fig time to pup}
p <- list_modify(
  p,
  time_pup =
    dfs_viz$long %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>% #ResToCsv("dev/ss/timeto")
    filter(instar == "pupa") %>%
    .plotfn() +
    labs(title = "hatch to pupation",
         )
)

p$time_pup
```

### supp
```{r supp fig time to pup}
ps <- list_modify(
  ps,
  time_pup = 
    dfs_viz$long %>%
    CalcDevSS_ad() %>% #ResToCsv("dev/ss/timeto-supp")
    filter(instar == "pupa") %>%
    .plotfn() +
    facet_wrap(~ year) +
    labs(y = "larval duration (days)")
)

ps$time_pup
```

### stats
```{r main stats pup dev rate}
mods <- list(
  m1 = lm(tt.pupa ~ year*pop*trt.minT,
           data = dfs_stats$dev_expt),
  m1a = lm(tt.pupa ~ year*pop*trt.minT + as.factor(is.sup),
           data = dfs_stats$dev_expt)
)

AIC(mods$m1, mods$m1a) # including sup is better lol
anova(mods$m1) # ok to focus on 2ways
anova(mods$m1a)

mods <- list_modify(
  mods,
  m2 = lm(tt.pupa ~ (year + pop + trt.minT)^2,
           dfs_stats$dev_expt)
)

AIC(mods$m2, mods$m1a)
anova(mods$m2) #%>% ResToCsv("dev/anova/timeto-pup")
summary(mods$m2)
```


## time to eclosion

outliers are only in the ctrl trt so just ignore ig

```{r main fig pup duration}
p <- list_modify(
  p,
  time_ec = 
    dfs_viz$long %>%
    filter(instar == "eclose",
           #tt < 45
           ) %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>% 
    # mutate(avg.tt = avg.tt - 15,
    #        se.tt = se.tt - 15) %>%
    .plotfn() +
    labs(title = "pupation to eclosion",
         )
)

p$time_ec
```

### supp

```{r}
ps <- list_modify(
  ps,
  time_ec = 
    dfs_viz$long %>%
    filter(instar == "eclose",
           #tt < 45
    ) %>%
    CalcDevSS_ad() %>%
    .plotfn() +
    facet_wrap(~ year) +
    labs(y = "pupal duration (days)")
)

ps$time_ec
```

### stats
```{r}
mods <- list(
  m1 = lm(tt.eclose ~ year*trt.minT*pop,
          data = dfs_stats$dev_expt)
)

# # look for outliers
# autoplot(mods$m1, which = c(4,6))
# 
# # high leverage
# dfs_stats$dev_expt[which(hatvalues(mods$m1) > 2*length(coef(mods$m1))/length(residuals(mods$m1))),] #%>% View()
# # high influence
# dfs_stats$dev_expt[which(cooks.distance(mods$m1) > 4/length(residuals(mods$m1))),] #%>% View()

anova(mods$m1) #%>% ResToCsv("dev/anova/timeto-ec") # keep the 3way
# i think this is prob bc of how Late the 2024 bugs were collected LOL
```


## collated figs

```{r}
# main
p$time_pup+ p$time_ec +
  plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5)) &
  ylim(c(20, 33))

#ResToFig("dev-time")

# supp
(ps$time_pup +
    theme(legend.position = "top",
          legend.margin = margin()) +
    labs(x = element_blank())) / 
  (ps$time_ec +
     theme(legend.position = "none") +
     guides(col = "none", lty = "none", shape = "none"))  +
  plot_layout(axes = "collect", #guides = "collect"
              ) 

#ResToFig("supp/dev-time")
```
## sup
```{r}
dfs_viz$long %>%
  filter(is.pup > 0) %>%
  group_by(is.sup) %>%
  CalcDevSS_ad() %>% 
  filter(instar == "pupa") %>% #View()
  .plotfn() +
  aes(alpha = as.factor(is.sup),
      group = interaction(trt.type, pop, is.sup)) +
  scale_alpha_manual(values = c(`0` = 1, `1` = 0.7)) +
  facet_grid(#is.sup
             ~ year)
```



# (4) longevity

## [>] longevity

```{r}
dfs_viz$long %>%
  filter(if.fridge == 0,
         instar == "long",
         if.mated == 1, 
         year != 2024) %>%
  #group_by(across(sex)) %>% # dont group by sex, per stats
  CalcDevSS_ad() %>% #View()
  Plot_Dev() +
  aes(y = avg.tt) +
  geom_point(size = 2) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.tt + se.tt,
                    ymin = avg.tt - se.tt),
                width = 0.25, lty = "solid", show.legend = FALSE) +
  #facet_grid(sex~year, labeller = labeller(sex = c(p_scales$facs_sex))) +
  facet_wrap(~ year) +
  labs(y = "lifespan (days)")
```

### [>] stats
do years separately!
```{r}
# df prep
dfs <- list(
  y23 = dfs_stats$dev_expt %>% filter(year == 2023),
  y25 = dfs_stats$dev_expt %>% filter(year == 2025)
) %>%
  lapply(., \(x){
    x %>%
      filter(if.fridge == 0, if.mated == 1)
  })
```

```{r}
# 2023 longev
mods <- list(
  m1 = lm(tt.long ~ pop*trt.minT, data = dfs$y23),
  m1a = lm(tt.long ~ pop*trt.minT*sex, data = dfs$y23)
)

anova(mods$m1, mods$m1a) # use the one w/o sex

mods <- list_modify(
  mods,
  m2 = lm(tt.long ~ pop*trt.minT + mass.eclose, data = dfs$y23)
)

anova(mods$m1, mods$m2) # ec mass doesnt improve model

anova(mods$m1) #%>% ResToCsv("dev/anova/long-23") # nothin matters!!
```

```{r}
# 2025 longev
mods <- list(
  m1 = lm(tt.long ~ pop*trt.minT, data = dfs$y25),
  m1a = lm(tt.long ~ pop*trt.minT*sex, data = dfs$y25)
)

anova(mods$m1, mods$m1a)  # use the one w/o sex

mods <- list_modify(
  mods,
  m2 = lm(tt.long ~ pop*trt.minT + mass.eclose, data = dfs$y25)
)

anova(mods$m1, mods$m2) # ec mass doesnt improve model

anova(mods$m1) #%>% ResToCsv("dev/anova/long-25") # minT matters...!!
```


## sample sizes
```{r sample sizes longevity, eval=FALSE}
dfs_viz$long %>%
  filter(if.fridge == 0,
         instar == "long",
         year != 2024,
         if.mated == 1) %>%
  # TODO include sex? see stats first
  CalcDevSS_ad() %>% 
  select(c(year, pop, trt.minT, trt.type, n,
           ends_with("tt"))) #%>% ResToCsv("dev/ss/timeto-longevity")
```

# ## (5) fitness

```{r}
dfs_viz$eggs %>%
    pivot_longer(cols = starts_with(c("avg", "se")),
               names_to = c(".value", "response"),
               names_pattern = "^(avg|se)\\.(.*)") %>%
  filter(response %in% c("sqrt.coll.perf", "p.hatch.perc")) %>%
  mutate(response = factor(response, levels = c("sqrt.coll.perf", "p.hatch.perc"))) %>%
  #ggplot(aes(y = avg, x = mate.trt)) +
  Plot_Fit() +
  geom_point() +
  geom_line() +
  geom_errorbar(aes(ymax = avg + se, ymin = avg - se),
                width = 0.25, lty = "solid", show.legend = FALSE) +
  aes(y = avg) +
  facet_grid(response ~ year, scales = "free_y",
             labeller = labeller(response = c("sqrt.coll.perf" = "sqrt[eggs collected\nper female]",
                                              "p.hatch.perc" = "proportion eggs hatched\nper enclosure")))
  

```

### stats (WIP)
```{r stats eggs laid}
# READING IN PROGRESS... see tents-stats.Rmd
mods <- list(
  all = lm(sqrt.coll.perf ~ n.f + mate.trt + mate.ishs + year,
           data = dfs$all)
)
```

### results table
```{r}
# TODO drop the colony F + expt M tents (from 2025?) but 
# this wasnt tracked in 2023...
dfs_viz$eggs %>%
  select(c(year, trt.minT, trt.mate, 
           starts_with(c("n.", "avg.", "se.")))) #%>% ResToCsv("egg results")
```

## sample sizes
```{r}
# TODO ideally use the same as the viz but maybe need to use the stats one lol cuz its on the per-tent level
# need to check the colony filtering on this...

# per cage
dfs_stats$eggs_all %>%
  filter(year != 2024) %>%
  group_by(year, trt.minT, trt.mate, cage) %>%
  summarise(perc.f = sum(n.f),
            perc.m = sum(n.m),
            perc.rat = n.f/n.m,
            perc.coll = sum(n.coll),
            perc.hatch = sum(n.hatch)) #%>% ResToCsv("tents/n/percage")

# per trt combo
dfs_stats$eggs_all %>%
  filter(year != 2024) %>% 
  mutate(rat = n.f/n.m,
         tot.ads = n.f + n.m) %>%
  group_by(year,  trt.minT, trt.mate) %>%
  summarise(n = n(),
            tot.f = sum(n.f),
            tot.m = sum(n.m),
            avg.rat = mean(rat),
            avg.ads = mean(tot.ads),
            tot.coll = sum(n.coll),
            tot.hatch = sum(n.hatch)) %>% View() #%>% ResToCsv("tents/n/pertrt")
```

# ## dump (ref only)

## supp dev patchwork figs

post-mortem: will use variable/aesthetic pairings that match what's being done in the main figs instead!


```{r eval = FALSE}
# add facet var
dfs_viz[c("wide", "long")] <- dfs_viz[c("wide", "long")] %>%
  lapply(., \(x){
    x %>%
      mutate(trt.facet = case_when(trt == 260 ~ "ctrl",
                                   pop == "lab" ~ "lab",
                                   pop == "field" ~ "field"),
             trt.facet = factor(trt.facet))
  })

# legend fn for patchwork plots (trt.type/yr)
AddLeg_Patchwork <- function(){
  list(
    labs(lty = "population", shape = "population"),
    scale_shape_manual(values = p_scales$shp_pop),
    scale_linetype_manual(values = p_scales$lty_pop),
    guides(shape = guide_legend(override.aes = list(linetype = c("255F", "solid"))))
  )
}
```

```{r supp fig pup survival patchwork, eval = FALSE}
.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = prop.pup,
               shape = pop, lty = pop)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    facet_wrap(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "% survived to pupation") +
    AddLeg_Patchwork() +
    scale_color_brewer(palette = "Dark2") +
    theme_bw()
}

p1 <- dfs_viz$wide %>%
  filter(trt == 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = year, 
      group = pop)

p2 <- dfs_viz$wide %>%
  filter(trt != 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = trt,
      group = year,
      col = year) +
  labs(x = p_scales$xlab_minT) +
  scale_x_discrete(labels = p_scales$labs_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0, 1))
```

```{r supp fig pup dev rate patchwork, eval = FALSE}
df <- dfs_viz$long %>%
  CalcDevSS_ad() %>%
  filter(instar == "pupa")

.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = avg.devrate,
               shape = pop, lty = pop
               )) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "larval development rate (1/days)") +
    scale_color_brewer(palette = "Dark2") +
    AddLeg_Patchwork() +
    theme_bw()
}

p1 <- df %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year, group = pop)

p2 <- df %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt, group = interaction(pop, year), col = year) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  labs(x = p_scales$xlab_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0.029, 0.054))
```

```{r supp fig adult longevity patchwork, eval = FALSE}
df <- dfs_viz$long %>%
    filter(if.fridge == 0,
           instar == "long") %>%
    CalcDevSS_ad()

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.tt,
               shape = pop, lty = pop
               )
           ) +
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.5, lty = "solid", show.legend = FALSE) +
    geom_point(size = 2) + 
    geom_line() +
    facet_grid(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "adult longevity (days)",) +
    theme_bw() +
    AddLeg_Patchwork() +
    ylim(c(4, 17.5))
}

p1 <-  df %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year,
      group = pop) +
  labs(x = "year")

p2 <- df %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt,
      col = year,
      group = interaction(year, pop)) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = p_scales$xlab_minT)

p1 + p2 +
  plot_layout(widths = c(1,2),
              axes = "collect", guides = "collect")
```