---
title: "final ntw results"
date: "2026-05-24"
---

# preamble and setup

```{r pathfinding, message=FALSE}
here::i_am("scripts/_mss/final-res.Rmd")
```

for collated viz + stats results for ntw mss

-   notes on the viz:
    -   main viz = focuses on trt type (NTW vs control), minimum temp, population
    -   supp viz = includes year and sex (if applicable)
-   notes on the stats:
    -   main stats = focuses on the NTW bugs and population, minimum temp as predictors
    - variable order does matter lol (ie `pop*minT` vs `minT*pop`). for the most part, will do models as pop first. 
    - using [this method](https://lmackerman.com/notebooks/prose_statistics.nb.html#Interaction_effect) for testing signif of effects (ie do LRTs via `anova(m1, m2)` instead of `drop1` (type ii anova/refits from dropping single terms) or `Anova()` (wald tests))
      - [read more](https://stats.stackexchange.com/questions/37805/why-do-anova-and-drop1-provided-different-answers-for-glmms)
    
main results/figs:

1. survival: hatch to pupation; hatch to eclosion
2. mass: pupal mass; eclosion mass
3. development time: hatch to pupation, pupation to eclosion
4. adult lifespan (2023 vs 2025)
5. adult reproduction (2023 vs 2025)

- everything should have a supp + stats section :)
- posthoc is in a separate script.


# loading

```{r data, message=FALSE}
source(here::here("scripts/_mss/load.R"))
```

```{r plot objects}
# for saving/exporting patchworks
p <- list() # main plots
ps <- list() # supp plots
```

# ## development

# (1) survival

## survival to pup

### main

```{r plot prep pup survival}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = prop.pup,
    ) +
    geom_line(show.legend = TRUE) +
    geom_point(size = 2,
               #aes(size = n)
               ) +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    #scale_radius(range = c(2,4)) +
    ylim(c(0, 1))
}
```

```{r main fig pup survival}
p <- list_modify(
  p,
  surv_pup =
    dfs_viz$wide %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps() %>% #ResToCsv("dev/ss/surv")
    .plotfn() +
    labs(#y = "% survived to pupation",
      title = "hatch to pupation"
    )
)

p$surv_pup
```

### supp
```{r supp fig pup survival}
ps <- list_modify(
  ps,
  surv_pup =
    dfs_viz$wide %>%
    CalcSurvProps_def() %>% #ResToCsv("dev/ss/surv-supp")
    .plotfn() +
    labs(y = "% survival to pupation") +
    facet_wrap(~year)
)

ps$surv_pup
```

### stats

```{r NEW main stats pup survival}
# fit models
mods <- list(
  int = glmer(is.pup ~ pop*trt.minT + (1|year),
              family = "binomial", data = dfs_stats$dev_expt),
  add = glmer(is.pup ~ pop + trt.minT + (1|year),
              family = "binomial", data = dfs_stats$dev_expt),
  no_pop = glmer(is.pup ~ trt.minT + (1|year),
                 family = "binomial", data = dfs_stats$dev_expt),
  no_minT = glmer(is.pup ~ pop + (1|year),
                  family = "binomial", data = dfs_stats$dev_expt),
  fixed = glm(is.pup ~ pop*trt.minT + year,
              family = "binomial", data = dfs_stats$dev_expt)
)

# # # test sig of effs (df, X^2, p) 
# anova(mods$int, mods$fixed)
 anova(mods$add, mods$int, test = "Chisq") # int: 2, 17.902,***

# anova(mods$no_pop, mods$add) # pop: 1, 42.646, ***
# anova(mods$no_minT, mods$add) # minT: 2, 78.428, ***

## tldr just do drop1 on the int and add models. (tho they lack the AIC/etc..)
## see: https://stats.stackexchange.com/questions/193643/likelihood-ratio-vs-score-vs-wald-test-different-p-values-which-to-use
## these refit the models and do LRT
drop1(mods$int, test = "Chisq") 
drop1(mods$add, test = "Chisq") 
# Anova(update(mods$int, 
#              contrasts = c("contr.sum", "contr.poly")), # should be list of x1 = contr, etc.
#       type = 3)

## (diff X^2 results from drop1 bc Anova does Wald tests (X^2) and not refitting model after dropping terms)
## also, cant specify test=F here.
# Anova(mods$int, type = 2)
# Anova(mods$add, type = 2)

# get effect sizes
anova(mods$int)
MuMIn::r.squaredGLMM(mods$int) # get R^2
summary(mods$int) 
mods$int %>% glance()
  #tidy() %>% write_clip() 
#%>% ResToCsv("summ/surv-pup") # missing var of rand effs tho
```


```{r eval = FALSE}
# trying afex
# library(afex)

## see .rds in data/ bc this took 5 yrs to compute
m <- mixed(is.pup ~ pop*trt.minT + (1|year),
           family = "binomial", data = dfs_stats$dev_expt,
           method = "PB")

summary(m) 
## main difference from below is in the ests for FEs and their p-vals

m1 <- mixed(is.pup ~ pop*trt.minT + (1|year),
           family = "binomial", data = dfs_stats$dev_expt,
           method = "LRT",
           #check_contrasts = FALSE # see note below
           )

summary(m1) 
## by default, mixed() uses sum-to-zero contrasts -- so will get different eff size estimates vs the glmer() default (treatment contrasts)
## see: https://github.com/singmann/afex/issues/4#issuecomment-145884270
```



```{r OLD main stats pup survival, eval = FALSE}
mods <- list(
  m1 = glm(is.pup ~ year*pop*trt.minT,
           family = "binomial", data = dfs_stats$dev_expt),
  m1a = glm(is.pup ~ year*pop*trt.minT + as.factor(is.sup),
           family = "binomial", data = dfs_stats$dev_expt)
)

AIC(mods$m1, mods$m1a) # technically better w sup but also lots of missing obs
anova(mods$m1, test = "Chisq") # can drop the 3-way

mods <- list_modify(
  mods,
  m2 = glm(is.pup ~ (year + pop + trt.minT)^2,
           family = "binomial", data = dfs_stats$dev_expt)
)

anova(mods$m2, test = "Chisq") #%>% ResToCsv("dev/anova/surv-pup")
summary(mods$m2)
```

## survival to eclosion

### main

```{r plot prep ec survival}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = prop.ec) +
    geom_line(show.legend = TRUE) +
    geom_point(size = 2) +
    geom_errorbar(aes(ymax = prop.ec + se.ec,
                      ymin = prop.ec - se.ec),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    ylim(c(0, 1))
}
```

```{r main fig adult survival}
# TODO include sample size?
p <- list_modify(
  p,
  surv_ec = 
    dfs_viz$wide %>%
    filter(is.na(is.LPI),
           #!is.na(sex) # can keep?
           ) %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
    CalcSurvProps() %>% 
    .plotfn() +
    labs(#y = "% pupa eclosed",
         title = "pupa to eclosion",
         )
)

p$surv_ec
```

### supp

```{r supp fig ec survival}
ps <- list_modify(
  ps,
  surv_ec = 
    dfs_viz$wide %>%
    CalcSurvProps_def() %>%
    .plotfn() +
    #geom_point(aes(size = n.pup)) +
    #geom_point(aes(size = sqrt(n))) +
    labs(y = "% survival to eclosion") +
    #scale_size_area(max_size = 6.5) +
    facet_wrap(~year)
  # TODO adjust y-lim if including pt size
)

ps$surv_ec
```

### stats

```{r NEW main stats ec survival}
# fit models
mods <- list(
  int = glmer(is.ec ~ pop*trt.minT + (1|year),
              family = "binomial", data = dfs_stats$dev_expt),
  ## singular fits
  add = glmer(is.ec ~ pop + trt.minT + (1|year),
              family = "binomial", data = dfs_stats$dev_expt),
  no_minT = glmer(is.ec ~ pop + (1|year),
              family = "binomial", data = dfs_stats$dev_expt),
  no_pop = glmer(is.ec ~ trt.minT + (1|year),
               family = "binomial", data = dfs_stats$dev_expt),
  ## glm versions
  add_F = glm(is.ec ~ pop + trt.minT + year,
              family = "binomial", data = dfs_stats$dev_expt),
  add_F2 = glm(is.ec ~ pop + trt.minT,
              family = "binomial", data = dfs_stats$dev_expt),
  int_F = glm(is.ec ~ pop*trt.minT,
              family = "binomial", data = dfs_stats$dev_expt)
)

# # test sig of effs (df, X^2, p)
anova(mods$int, mods$add) # int: 2, 2.488, 0.2882
# anova(mods$add,mods$no_minT) # minT: 2, 2.6909, 0.2604
# anova(mods$add, mods$no_pop) # pop: 1, 12.289, ***

drop1(mods$int, test = "Chisq") 
drop1(mods$add, test = "Chisq")
drop1(mods$add_F, test = "Chisq") # glm_add similar results to glmer_add

anova(mods$int, mods$int_F, mods$add_F2) 
# F2 is best, tho they are all similar. just use the glmer LOL

## comparing the additive models
summary(mods$add)
summary(mods$add_F) # similar betas to the glmer

# get effect sizes
summary(mods$int) # do on the int (instead add) bc otherwise var is actually 0 lol
# (bc of singular: https://cran.r-project.org/web/packages/afex/vignettes/afex_mixed_example.html#model-setup)
```

```{r OLD main stats surv to ec, eval = FALSE}
# filter df bc of including mass.pupa in the model -- m1
# will fit extra observations
df <- dfs_stats$dev_expt %>%
  filter(!is.na(mass.pupa))

mods <- list(
  m1 = glm(is.ec ~ year*pop*trt.minT,
           family = "binomial", data = df),
  m1a = glm(is.ec ~ year*pop*trt.minT + mass.pupa,
           family = "binomial", data = df)
)


drop1(mods$m1, scope = ~year + pop + trt.minT, test = "LRT")
drop1(mods$m1, test = "LRT")
drop1(mods$m1, .~., test = "LRT")

drop1(mods$m1a, test = "LRT")
# can drop the 3way...

AIC(mods$m1, mods$m1a) 
anova(mods$m1, mods$m1a, test = "Chisq")
# including mass is better, per anova + aic

#anova(mods$m1, test = "Chisq") # just year and pop matter lmao
anova(mods$m1a, test = "Chisq") # 3 way is marginally signif

# try to simplify?
mods <- list_modify(
  mods,
  m2 = glm(is.ec ~ (year + pop + trt.minT)^2 + mass.pupa,
           family = "binomial", data = df),
)

AIC(mods$m2, mods$m1a) 
anova(mods$m2, mods$m1a, test = "Chisq")
# more complex model is slightly better, per AIC

anova(mods$m1a, test = "Chisq") #%>% ResToCsv("dev/anova/surv-ec")
summary(mods$m1a)
```

```{r alt stats surv to ec}
# excluding pupal mass

mods <- list(
  m1 = glm(is.ec ~ year*pop*trt.minT,
           family = "binomial", data = dfs_stats$dev_expt)
)

drop1(mods$m1, scope = ~year + pop + trt.minT, test = "LRT")
drop1(mods$m1, .~., test = "LRT")

drop1(mods$m1, test = "LRT") # can drop the 3way...
anova(mods$m1, test = "Chisq") # 3way is ns

# 2 ways only
mods <- list_modify(
  mods,
  m2 = glm(is.ec ~ (year + pop + trt.minT)^2,
           family = "binomial", data = dfs_stats$dev_expt),
)

AIC(mods$m2, mods$m1)  # more complex is very slightly better
anova(mods$m2, mods$m1, test = "Chisq")
# tho not significantly different

anova(mods$m2, test = "Chisq") #%>% ResToCsv("dev/anova/surv-ec")
summary(mods$m2) # sth about 2024...
```

```{r}
# looking at pup mass
dfs_viz$wide %>%
    filter(is.na(is.LPI),
           !is.na(is.ec)
           ) %>%
    PrepSurvProps() %>%
    ungroup(year) %>%
  group_by(is.ec, .add = TRUE) %>%
  summarise(n = n(),
            avg.wmass = mean(mass.wander, na.rm = TRUE),
            se.wmass= se(mass.wander),
            avg.pmass = mean(mass.pupa, na.rm = TRUE),
            se.pmass = se(mass.pupa)) %>% View()
```


## collated figs

```{r}
# main
p$surv_pup + theme(legend.position = "none") + 
  p$surv_ec +
  plot_layout(axes = "collect", guides = "collect") &
  labs(y = "% survival") &
  theme(plot.title = element_text(hjust = 0.5))  

#ResToFig("dev-surv")

# supp
(ps$surv_pup + 
    theme(legend.position = "top",
                     legend.margin = margin()) +
    labs(y = "% survival\nto pupation",
                    x = element_blank())
  ) / 
  (ps$surv_ec + 
     labs(y = "% survival\nto eclosion") +
     guides(col = "none", shape = "none", lty = "none") +
     theme(legend.position = "none")
   ) +
  plot_layout(axes = "collect", #guides = "collect"
              )

#ResToFig("supp/dev-surv")
```

## sample sizes
```{r}
dfs_viz$wide %>%
  group_by(year, pop, trt) %>%
  summarise(n = sum(!is.na(is.pup)),
            wand = sum(!is.na(jdate.wander), na.rm = TRUE),
            pup = sum(is.pup == 1, na.rm = TRUE),
            ec = sum(is.ec == 1, na.rm = TRUE),
            LPI = sum(is.pup == 2, na.rm = TRUE)) #%>% ResToCsv("dev/n/surv")
```


# (2) mass 

```{r plot fn mass}
.plotfn <- function(x) {
  x %>%
    Plot_Dev() +
    aes(y = avg.mass) +
    geom_point(size = 2) +
    geom_line(show.legend = TRUE) +
    geom_errorbar(aes(ymax = avg.mass + se.mass,
                      ymin = avg.mass - se.mass),
                  width = 0.5, lty = "solid", show.legend = FALSE,
                  ) +
    labs(y = "mass (g)")
}
```

## pupal mass

### main

drop LPIs bc mass isnt accurate

```{r df prep pup mass}
df <- dfs_viz$long %>%
  filter(!is.na(sex),
         instar == "pupa")
```

```{r main figs pup mass}
p <- list_modify(
  p,
  mass_pup = # sexes combined
    df %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    .plotfn() +
    labs(title = "pupa"),
  
  mass_pupsex = # sexes parsed
    df %>%
    PrepAdultSS() %>%
    group_by(sex, .add = TRUE) %>%
    ungroup(year) %>%
    CalcDevSS() %>% #ResToCsv("dev/ss/mass-pup-sex")
    .plotfn() +
    facet_wrap(~ sex, labeller = as_labeller(p_scales$facs_sex)) +
    labs(title = "pupa") 
)

p$mass_pup
p$mass_pupsex # use this, per stats
```

### supp

```{r supp figs pup mass}
ps <- list_modify(
  ps,
  # sexes combined
  mass_pup =
    df %>%
    PrepAdultSS() %>%
    ungroup(starts_with("trt")) %>% 
    group_by(trt, trt.type, trt.minT, .add = TRUE) %>%
    CalcDevSS() %>% #View()
    .plotfn() +
    facet_wrap(~ year) +
    ylab("pupal mass (g)"),
  
  # sexes parsed (use this one)
  mass_pupsex =
  df %>%
    PrepAdultSS() %>%
    ungroup(starts_with("trt")) %>% 
    group_by(trt, trt.type, trt.minT, sex, .add = TRUE) %>% 
    CalcDevSS() %>% #ResToCsv("dev/ss/mass-pup-sex-supp") 
    #filter(trt != 260) %>%
    .plotfn() +
    facet_grid(sex ~ year, labeller = labeller(sex = c(p_scales$facs_sex)),
               #scales = "free"
               ) +
    ylab("pupal mass (g)")
)

ps$mass_pup
ps$mass_pupsex
```


### stats

```{r NEW main stats pup mass}
# fit models
df <- dfs_stats$dev_expt %>% filter(!is.na(sex))

mods <- list(
  int = lmer(mass.pupa ~ pop*trt.minT + sex + (1|year),
             data = df),
  int2 = lmer(mass.pupa ~ pop*trt.minT + (1|year),
             data = df)
)

## include sex? -- yes
## (variations of tests...)
#anova(mods$int2, mods$int) # sex: 1, 6.3869, *
drop1(mods$int, scope = c("sex")) # keep sex
# drop1(mods$int, scope = c("sex"), ddf = "lme4", test = "Chisq")
# drop1(mods$int, scope = c("sex", "pop:trt.minT"), test = "Chisq")

# fit full and depleted models
mods <- list_modify(
  mods,
  add = lmer(mass.pupa ~ pop + trt.minT + sex + (1|year),
             data = df),
  no_minT = lmer(mass.pupa ~ pop + sex + (1|year),
             data = df),
  no_pop = lmer(mass.pupa ~ trt.minT + sex + (1|year),
             data = df)
  # no_sex mod is just the above comparison
)

# # test sig of effs (df, X^2, p)
# anova(mods$int, mods$add) %>% write_clip() # int: 2, 8.4835, *
# drop1(mods$int, scope = c("pop:trt.minT"), test = "Chisq")
# drop1(mods$int, scope = c("pop:trt.minT"), ddf = "lme4", test = "Chisq")
# 
# anova(mods$add, mods$no_pop) # pop: 1, 126.32, ***
 anova(mods$add,mods$no_minT, test = "F") # minT: 2, 36.642, ***
# drop1(mods$add, scope = c("pop", "trt.minT", "sex"), test = "Chisq")
# drop1(mods$add, scope = c("pop", "trt.minT", "sex"), ddf = "lme4", test = "Chisq")

# drop1(mods$int, scope = c("pop:trt.minT")) 
# drop1(mods$add)
 drop1(lm(mass.pupa ~ pop*trt.minT + sex + year, data = df), 
       scope = ~ .)
# jk just use anova.lmerModLmerTest
anova(mods$int, type = 2) # F test
#Anova(mods$int, type = 2) # Wald X^2

# get effect sizes
summary(mods$int)
```

```{r OLD main stats pup mass, eval = FALSE}
mods <- list(
  m1 = lm(mass.pupa ~ year*pop*trt.minT*sex,
           data = dfs_stats$dev_expt
          )
)

anova(mods$m1) # ok to just focus on the 2 ways of trt/pop/year and main effect of sex

mods <- list_modify(
  mods,
  m2 = lm(mass.pupa ~ (year + pop + trt.minT + sex)^2,
           data = dfs_stats$dev_expt),
  m3 = lm(mass.pupa ~ (year + pop + trt.minT)^2 + sex,
           data = dfs_stats$dev_expt)
)

anova(mods$m3, mods$m2, mods$m1)
## adding model complexity doesnt do much
AIC(mods$m3, mods$m2, mods$m1)
## m3 = lowest AIC

anova(mods$m3) #%>% ResToCsv("dev/anova/mass-pup")
summary(mods$m3)
```

## eclosion mass

### main

```{r df prep eclosion mass}
df <- dfs_viz$long %>%
  filter(!is.na(sex),
         instar == "eclose")
```

```{r main fig eclosion mass}
p <- list_modify(
  p,
  mass_ec = # sexes combined
    df %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>%
    .plotfn() +
    labs(title = "adult"),
  
  mass_ecsex = # sexes parsed
    df %>%
    PrepAdultSS() %>%
    group_by(sex, .add = TRUE) %>%
    ungroup(year) %>%
    CalcDevSS() %>% #ResToCsv("dev/ss/mass-ec-sex")
    .plotfn() +
    facet_wrap(~ sex, labeller = as_labeller(p_scales$facs_sex)) +
    labs(title = "adult")
)

p$mass_ec
p$mass_ecsex
```

### supp

```{r supp figs eclosion mass}
ps <- list_modify(
  ps,
  # sexes combined
  mass_ec =
    df %>%
    #filter(trt.type == "expt") %>%
    PrepAdultSS() %>%
    ungroup(starts_with("trt")) %>% 
    group_by(trt, trt.type, trt.minT, .add = TRUE) %>%
    CalcDevSS() %>%
    .plotfn() +
    #geom_point(aes(size = n)) +
    facet_wrap(~ year) +
    ylab("adult mass (g)"),
  
  # sexes parsed 
  mass_ecsex =
    df %>%
    #filter(trt.type == "expt") %>%
    PrepAdultSS() %>%
    ungroup(starts_with("trt")) %>% 
    group_by(trt, trt.type, trt.minT, sex, .add = TRUE) %>%
    CalcDevSS() %>% #ResToCsv("dev/ss/mass-ec-sex-supp")
    #filter(trt != 260) %>%
    .plotfn() +
    #geom_point(aes(size = n)) +
    facet_grid(sex ~ year, labeller = labeller(sex = c(p_scales$facs_sex)),
               #scales = "free"
               ) +
    ylab("adult mass (g)")
)

ps$mass_ec
ps$mass_ecsex
```

### stats

```{r NEW main stats eclosion mass}
# fit models
df <- dfs_stats$dev_expt %>% filter(!is.na(sex))

mods <- list(
  int = lmer(mass.eclose ~ pop*trt.minT + sex + (1|year),
             data = df),
  int2 = lmer(mass.eclose ~ pop*trt.minT + (1|year),
             data = df)
)

## include sex? -- yes
#anova(mods$int2, mods$int) # sex: 1, 27.7, ***
drop1(mods$int, scope = "sex") #yes

# fit full and depleted models
mods <- list_modify(
  mods,
  add = lmer(mass.eclose ~ pop + trt.minT + sex + (1|year),
             data = df),
  no_minT = lmer(mass.eclose ~ pop + sex + (1|year),
             data = df),
  no_pop = lmer(mass.eclose ~ trt.minT + sex + (1|year),
             data = df)
  # no_sex mod is just the above comparison
)

# # test sig of effs (df, X^2, p)
# anova(mods$int, mods$add) # int: 2, 10.683, **
# anova(mods$add, mods$no_pop) # pop: 1, 40.148, ***
# anova(mods$add,mods$no_minT) # minT: 2, 7.6602, *

# drop1(mods$int, scope = "pop:trt.minT")
# drop1(mods$add)
anova(mods$int, type = 2) 

# get effect sizes
summary(mods$int)
```

```{r OLD eclosion stats, eval = FALSE}
# include pupal mass?
mods <- list(
    m1 = lm(mass.eclose ~ year*pop*trt.minT*sex + mass.pupa,
           data = dfs_stats$dev_expt),
    m1a = lm(mass.eclose ~ year*pop*trt.minT*sex,
            data = dfs_stats$dev_expt),
    m1b = lm(mass.eclose ~ year*pop*trt.minT + mass.pupa,
            data = dfs_stats$dev_expt)
)

AIC(mods$m1, mods$m1a, mods$m1b) # yes, based on AIC

# which interactions to keep?
anova(mods$m1) # main effs, and 2-ways with minT:sex, year:sex

mods <- list_modify(
  mods,
  m3 = lm(mass.eclose ~ (year + pop + trt.minT + sex)^2 + mass.pupa,
          data = dfs_stats$dev_expt),
  m4 = lm(mass.eclose ~ (year + pop + trt.minT)^2 + sex + mass.pupa,
          data = dfs_stats$dev_expt),
  m5 = lm(mass.eclose ~ year*sex + pop + trt.minT*sex + mass.pupa,
          data = dfs_stats$dev_expt)
)

anova(mods$m3)
anova(mods$m4)
AIC(mods$m1, mods$m3, mods$m4, mods$m5) # isolating the 2ways is better

anova(mods$m3, mods$m4) #m4 better
anova(mods$m4, mods$m5) #m5 not better than m4?
anova(mods$m3, mods$m4, mods$m5)

# just keep the one with the 2 ways bc way easier to explain in methods lol
anova(mods$m3) #%>% ResToCsv("dev/anova/mass-ec")
summary(mods$m3)
```



## collated figs

```{r}

# main
p$mass_pupsex + theme(legend.position = "none") +
  p$mass_ecsex +
  plot_layout(axes = "collect", #guides = "collect"
              widths = c(0.9, 1)) &
  labs(x = p_scales$xlab_minT) &
  theme(plot.title = element_text(hjust = 0.5))

#ResToFig("dev-mass")

# supp
(ps$mass_pupsex +
    theme(legend.position = "top",
          legend.margin = margin()) +
    labs(x = element_blank())) / 
  (ps$mass_ecsex +
     theme(legend.position = "none") +
     guides(col = "none", lty = "none", shape = "none"))  +
  plot_layout(axes = "collect", #guides = "collect"
              )

#ResToFig("supp/dev-mass")
```

# (3) development time

```{r plot prep dev time}
.plotfn <- function(x){
  x %>%
    Plot_Dev() +
    aes(y = avg.tt) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.tt + se.tt,
                      ymin = avg.tt - se.tt),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "duration (days)")
}
```

## time to pupation

### main

```{r main fig time to pup}
p <- list_modify(
  p,
  time_pup =
    dfs_viz$long %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>% #ResToCsv("dev/ss/timeto")
    filter(instar == "pupa") %>%
    .plotfn() +
    labs(title = "hatch to pupation",
         )
)

p$time_pup
```

### supp
```{r supp fig time to pup}
ps <- list_modify(
  ps,
  time_pup = 
    dfs_viz$long %>%
    CalcDevSS_ad() %>% #ResToCsv("dev/ss/timeto-supp")
    filter(instar == "pupa") %>%
    .plotfn() +
    facet_wrap(~ year) +
    labs(y = "larval duration (days)")
)

ps$time_pup
```

### stats

```{r NEW main stats pup time}
# fit full and depleted models
mods <- list(
  int = lmer(tt.pupa ~ pop*trt.minT + (1|year),
             data = dfs_stats$dev_expt),
  add = lmer(tt.pupa ~ pop + trt.minT + (1|year),
             data = dfs_stats$dev_expt),
  no_minT = lmer(tt.pupa ~ pop + (1|year),
             data = dfs_stats$dev_expt),
  no_pop = lmer(tt.pupa ~ trt.minT + (1|year),
             data = dfs_stats$dev_expt)
)

# # test sig of effs (df, X^2, p)
# anova(mods$int, mods$add) # int: 2, 2.6775, 0.2622
# anova(mods$add, mods$no_pop) # pop: 1, 22.863, ***
# anova(mods$add,mods$no_minT) # minT: 2, 133.33, ***

# drop1(mods$int) # int is ns
# drop1(mods$add) 
anova(mods$int, type = 2)

# get effect sizes 
# TODO: revist. they have diff results
summary(mods$int)
summary(mods$add)
```

```{r OLD main stats pup dev rate, eval = FALSE}
mods <- list(
  m1 = lm(tt.pupa ~ year*pop*trt.minT,
           data = dfs_stats$dev_expt),
  m1a = lm(tt.pupa ~ year*pop*trt.minT + as.factor(is.sup),
           data = dfs_stats$dev_expt)
)

AIC(mods$m1, mods$m1a) # including sup is better lol
anova(mods$m1) # ok to focus on 2ways
anova(mods$m1a)

mods <- list_modify(
  mods,
  m2 = lm(tt.pupa ~ (year + pop + trt.minT)^2,
           dfs_stats$dev_expt),
  m2a = lm(tt.pupa ~ (year + pop + trt.minT)^2 + as.factor(is.sup),
           dfs_stats$dev_expt)
)

AIC(mods$m2, mods$m2a, mods$m1, mods$m1a)
anova(mods$m2) #%>% ResToCsv("dev/anova/timeto-pup")
summary(mods$m2)
```


## time to eclosion

### main

outliers are only in the ctrl trt so just ignore ig

```{r main fig pup duration}
p <- list_modify(
  p,
  time_ec = 
    dfs_viz$long %>%
    filter(instar == "eclose",
           #tt < 45
           ) %>%
    PrepAdultSS() %>%
    ungroup(year) %>%
    CalcDevSS() %>% 
    # mutate(avg.tt = avg.tt - 15,
    #        se.tt = se.tt - 15) %>%
    .plotfn() +
    labs(title = "pupation to eclosion",
         )
)

p$time_ec
```

### supp

```{r}
ps <- list_modify(
  ps,
  time_ec = 
    dfs_viz$long %>%
    filter(instar == "eclose",
           #tt < 45
    ) %>%
    CalcDevSS_ad() %>%
    .plotfn() +
    facet_wrap(~ year) +
    labs(y = "pupal duration (days)")
)

ps$time_ec
```

### stats
```{r NEW main stats ec time}
# fit full and depleted models
mods <- list(
  int = lmer(tt.eclose ~ pop*trt.minT + (1|year),
             data = dfs_stats$dev_expt),
  add = lmer(tt.eclose ~ pop + trt.minT + (1|year),
             data = dfs_stats$dev_expt),
  no_minT = lmer(tt.eclose ~ pop + (1|year),
             data = dfs_stats$dev_expt),
  no_pop = lmer(tt.eclose ~ trt.minT + (1|year),
             data = dfs_stats$dev_expt)
)

# # test sig of effs (df, X^2, p)
# Anova(mods$int, test.statistic="Chisq")
# Anova(mods$int, test.statistic="F")
# drop1(mods$int, ddf = "Kenward-Roger")
# drop1(mods$int)
# 
# anova(mods$int, mods$add) # int: 2, 3.3228, 0.1899
# anova(mods$add, mods$no_pop) # pop: 1, 0.8621, 0.3532
# anova(mods$add,mods$no_minT) # minT: 2, 41.554, ***

# drop1(mods$int) # ns
# drop1(mods$add)
anova(mods$int, type = 2) 


# get effect sizes 
# TODO: revist. results are same but idk which to use
# see: https://stats.stackexchange.com/questions/469329/intercept-significant-in-glmer-and-lmer#469354
summary(mods$int)
summary(mods$add)
```

```{r OLD stats ec time, eval = FALSE}
mods <- list(
  m1 = lm(tt.eclose ~ year*trt.minT*pop,
          data = dfs_stats$dev_expt)
)

# # look for outliers
# autoplot(mods$m1, which = c(4,6))
# 
# # high leverage
# dfs_stats$dev_expt[which(hatvalues(mods$m1) > 2*length(coef(mods$m1))/length(residuals(mods$m1))),] #%>% View()
# # high influence
# dfs_stats$dev_expt[which(cooks.distance(mods$m1) > 4/length(residuals(mods$m1))),] #%>% View()

anova(mods$m1) #%>% ResToCsv("dev/anova/timeto-ec") # keep the 3way
# i think this is prob bc of how Late the 2024 bugs were collected LOL
```


## collated figs

```{r}
# main
p$time_pup+ p$time_ec +
  plot_layout(axes = "collect", guides = "collect") &
  theme(plot.title = element_text(hjust = 0.5)) &
  ylim(c(20, 33))

#ResToFig("dev-time")

# supp
(ps$time_pup +
    theme(legend.position = "top",
          legend.margin = margin()) +
    labs(x = element_blank())) / 
  (ps$time_ec +
     theme(legend.position = "none") +
     guides(col = "none", lty = "none", shape = "none"))  +
  plot_layout(axes = "collect", #guides = "collect"
              ) 

#ResToFig("supp/dev-time")
```

## supernumerary

```{r}
dfs_viz$long %>%
  filter(is.pup > 0,
         !is.na(pop)) %>%
  group_by(is.sup) %>%
  CalcDevSS_ad() %>% 
  filter(instar == "pupa") %>% #View()
  .plotfn() +
  #aes(alpha = as.factor(is.sup)) +
  geom_line(aes(group = interaction(trt.type, pop))) +
  #scale_alpha_manual(values = c(`0` = 1, `1` = 0.7)) +
  facet_grid(is.sup ~ year, labeller = labeller(is.sup = c("0" = "5 instars",
                                                           "1" = "6+ instars"))) +
  labs(y = "time to pupation (days)")
```

# (4) lifespan

- comments on EDAs:
  - mating status matters, so only use mated bugs.
  - doesnt seem to be affected if you're a colony bug.
  - effect of fridge: not signif if yrs are consolidated for mated bugs only. if yrs are parsed, signif in 2025 only but probably bc of N.
  - effect of sex: only matters for expt, mated bugs (if yrs are consolidated)
## lifespan

### main

```{r}
dfs_viz$long %>%
  filter(if.fridge == 0,
         instar == "long",
         if.mated == 1, 
         year != 2024) %>%
  #group_by(across(sex)) %>% # dont group by sex, per stats
  CalcDevSS_ad() %>% #View()
  Plot_Dev() +
  aes(y = avg.tt) +
  geom_point(aes(size = n),
             alpha = 0.6) +
  geom_line() +
  geom_errorbar(aes(ymax = avg.tt + se.tt,
                    ymin = avg.tt - se.tt),
                width = 0.25, lty = "solid", show.legend = FALSE) +
  #facet_grid(sex~year, labeller = labeller(sex = c(p_scales$facs_sex))) +
  facet_wrap(~ year) +
  ylim(c(2, 15)) +
  labs(y = "lifespan (days)",
       size = "N") +
  scale_size_area(max_size = 6.5) 
  
#ResToFig("dev-long_nofridge")
```


### stats

dropping fridge

```{r}
# fridge effect
df <- dfs_stats$dev_expt %>%
  filter(if.mated == 1)

mods <- list(
  all23 = lm(tt.long ~ pop*trt.minT + if.fridge, 
             data = df %>% filter(year == 2023)),
  nof23 = lm(tt.long ~ pop*trt.minT, 
             data = df %>% filter(year == 2023,
                                  if.fridge == 0)),
  all25 = lm(tt.long ~ pop*trt.minT + if.fridge, 
             data = df %>% filter(year == 2025)),
  nof25 = lm(tt.long ~ pop*trt.minT, 
             data = df %>% filter(year == 2025,
                                  if.fridge == 0))
)

Anova(mods$all23, type = 2)
Anova(mods$nof23, type = 2)
## doesnt matter for 2023 as much

Anova(mods$all25, type = 2)
Anova(mods$nof25, type = 2)
## matters in 2025 but results are still similar if fridge is dropped
```



```{r}
# 2023
mods <- lm(tt.long ~ pop*trt.minT, 
             data = dfs_stats$dev_expt %>% filter(year == 2023,
                                                 if.mated == 1,
                                                 if.fridge == 0))

drop1(mods, scope = ~ ., test = "F") %>% write_clip()
# nothing is sig
```

```{r}
# 2025
mods <- lm(tt.long ~ pop*trt.minT, 
             data = dfs_stats$dev_expt %>% filter(year == 2025,
                                                 if.mated == 1,
                                                 if.fridge == 0))

# not sure why this isnt the same... cuz drop1 doesnt refit?
anova(update(mods, ~ . - pop:trt.minT),
      update(mods, ~ trt.minT), test = "F")

drop1(mods, scope = ~ ., test = "F")
# trt is sig
```



### stats (old)
do years separately!

```{r}
# df prep
dfs <- list(
  all = dfs_stats$dev_all,
  y23 = dfs_stats$dev_expt %>% filter(year == 2023),
  y25 = dfs_stats$dev_expt %>% filter(year == 2025)
) %>%
  lapply(., \(x){
    x %>%
      filter(#if.fridge == 0, 
             if.mated == 1)
  })
```

```{r}
# 2023 longev

# check if ctrls and expt are different
anova(lm(tt.long ~ pop*trt.type, data = filter(dfs$all, year == 2023)))
anova(lm(tt.long ~ pop*trt, data = filter(dfs$all, year == 2023)))
## nope!

# ANOVA
mods <- list(
  m1 = lm(tt.long ~ pop*trt.minT, data = dfs$y23),
  m1a = lm(tt.long ~ pop*trt.minT*sex, data = dfs$y23)
)

add1(mods$m1, scope = ~ . + sex + mass.eclose, test = "F")

#add1(mods$m1a, scope = formula(mods$m1a)[-2], test = "F")
anova(mods$m1, mods$m1a) # sex does not improve model

mods <- list_modify(
  mods,
  m2 = lm(tt.long ~ pop*trt.minT + mass.eclose, data = dfs$y23)
)

anova(mods$m1, mods$m2) # ec mass doesnt improve model

anova(mods$m1) #%>% ResToCsv("dev/anova/long-23") # nothin matters!!
```

```{r}
# 2025 longev

# check if ctrls and expt are different
anova(lm(tt.long ~ pop*trt.type, data = filter(dfs$all, year == 2025)))
anova(lm(tt.long ~ pop*trt, data = filter(dfs$all, year == 2025))) # yes, but prob bc of unbalanced groups
## so nope!

# ANOVA
mods <- list(
  m1 = lm(tt.long ~ pop*trt.minT, data = dfs$y25),
  m1a = lm(tt.long ~ pop*trt.minT*sex, data = dfs$y25)
)

add1(mods$m1, scope = ~ . + sex + mass.eclose, test = "F")
anova(mods$m1, mods$m1a)  # sex doesnt improve model

mods <- list_modify(
  mods,
  m2 = lm(tt.long ~ pop*trt.minT + mass.eclose, data = dfs$y25)
)

anova(mods$m1, mods$m2) # if no fridge bugs, ec mass doesnt improve model. otherwise marginal

anova(mods$m1) #%>% ResToCsv("dev/anova/long-25") # minT matters...!!

# if you drop 40-33?
anova(lm(tt.long ~ pop*trt.minT, data = dfs$y25 %>% filter(trt.minT != 33))) # similar results
```

## sample sizes
```{r sample sizes longevity, eval=FALSE}
dfs_viz$long %>%
  filter(if.fridge == 0,
         instar == "long",
         year != 2024,
         if.mated == 1) %>%
  CalcDevSS_ad() %>% 
  select(c(year, pop, trt, n,
           ends_with("tt"))) #%>% ResToCsv("dev/ss/timeto-longevity")
```

# ## (5) reproduction
# reproduction
TODO
- check that all the tents being counted are actually tents for mating lol (e.g. wtf is 107-H in 2023)

## eggs collected

### main

```{r box and whisker}
pd <- 0.75

gg <- dfs_viz$eggs_n %>%
  ggplot(aes(y = n.coll.perf)) +
  geom_boxplot(position = position_dodge(width = pd),
               outlier.shape = NA) +
  geom_point(data = dfs_viz$eggs_n,
             aes(size = n.f),
             position = position_dodge(width = pd),
             alpha = 0.5) +
  #scale_size(range = c(1.5, 5.5)) +
  scale_size_area(max_size = 6) +
  labs(y = "total eggs collected per female",
       size = "total females",
       ) +
  facet_wrap(~year) +
  theme_bw() +
  theme(legend.position = "top",
        legend.box = "vertical",
        legend.margin = margin()) +
  guides(color = guide_legend(order = 1),
         size = guide_legend(order = 2))

# x as minT
gg +
  aes(x = trt.minT, col = trt.mate,) +
  #geom_point(aes(group = trt.mate)) +
  scale_color_manual(values = p_scales$cols_trtsex, labels = p_scales$labs_trtmate) +
  scale_x_discrete(labels = p_scales$labs_trt) +
  labs(col = "heat-treated parent",
       x = p_scales$xlab_minT)
#ResToFig("fit-coll-bytrt")

# x as trted sex
gg + 
  aes(col = trt.minT, x = trt.mate) +
  #geom_point(aes(group = trt.minT)) +
  scale_color_manual(values = p_scales$cols_minT, labels = p_scales$labs_minT) +
  scale_x_discrete(labels = p_scales$labs_trtmate) +
  labs(col = p_scales$xlab_minT,
       x = "heat-treated parent")
#ResToFig("fit-coll-bysex")
```



```{r point plot, eval=FALSE}
.plotfn <- function(x){
  x %>%
    Plot_Fit() +
    geom_point() +
    geom_line() +
    geom_errorbar(aes(ymax = avg + se, ymin = avg - se),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    aes(y = avg) +
    facet_wrap(~ year) 
}

p <- list_modify(
  p,
  fit_coll = dfs_viz$eggs_ss %>%
    SSToLong() %>%
    filter(response == "sqrt.coll.perf") %>%
    .plotfn() +
    labs(y = "sqrt[eggs collected\nper female]"),
  
  fit_hatch = dfs_viz$eggs_ss %>%
  SSToLong() %>%
  filter(response == "p.hatch.perc") %>%
  .plotfn() +
  labs(y = "proportion eggs hatched\nper enclosure") 
) 

p$fit_coll
p$fit_hatch

(p$fit_coll +
    theme(strip.background.y = element_blank(),
          strip.text.y = element_blank(),
          axis.title.x = element_blank()
          )
  ) /
  (p$fit_hatch +
     theme(strip.background = element_blank(),
           strip.text = element_blank(),
           legend.position = "none"
           )
   ) +
  plot_layout(axes = "collect", guides = "collect"
              ) &
  theme(legend.position = "top",
        axis.title.y = element_text(margin = margin(), size = 10))
#ResToFig("fit-pts")

```

### stats

```{r stats eggs laid 2023}
# 2023
df <- dfs_stats$eggs %>%
  filter(year == 2023)

mods <- list(
  all_win = glmer(n.coll ~ trt.type + (1|cage), offset = log(n.f),
                  family = "poisson", data = df %>% filter(mate.type == "within")),
  expt_win = glmer(n.coll ~ trt.minT + (1|cage), offset = log(n.f),
                   family = "poisson", data = df %>% filter(mate.type == "within" & trt.type == "expt")),
  
  # does not converge if int
  all = glmer(n.coll ~ trt.minT * mate.ishs + (1|cage), offset = log(n.f),
              family = "poisson", data = df),
  all2 = glmer(n.coll ~ trt.minT + mate.ishs + (1|cage), offset = log(n.f),
              family = "poisson", data = df),
  expt_btwn = glmer(n.coll ~ trt.minT*mate.ishs + (1|cage), offset = log(n.f),
                    family = "poisson", data = df %>% filter(mate.type == "between")),
  expt_all = glmer(n.coll ~ trt.minT * mate.ishs + (1|cage), offset = log(n.f),
                   family = "poisson", data = df %>% filter(trt.type == "expt"))
)

# trt effects
drop1(mods$all_win, test = "Chisq") # no diff btwn ctrl/hs LOLOL
drop1(mods$expt_win, test = "Chisq") # no diff btwn hs groups

# sex effects
#drop1(mods$all, test = "Chisq") # int is ns
drop1(update(mods$all, ~ . - trt.minT:mate.ishs), test = "Chisq") # no eff of minT or if m/f/both/none
drop1(mods$all2, test = "Chisq")

#drop1(mods$expt_btwn, test = "Chisq") # negligible int
drop1(update(mods$expt_btwn, ~ . - trt.minT:mate.ishs), test = "Chisq") # no main effects

#drop1(mods$expt_all, test = "Chisq") # interaction is ns
drop1(update(mods$expt_all, ~ . - trt.minT:mate.ishs), test = "Chisq") # no effect of minT or hs!
```

```{r stats eggs laid 2025}
# 2025
df <- dfs_stats$eggs %>%
  filter(year == 2025)

mods <- list(
  all_win = glmer(n.coll ~ trt.type + (1|cage), offset = log(n.f),
                  family = "poisson", data = df %>% filter(mate.type == "within")),
  expt_win = glmer(n.coll ~ trt.minT + (1|cage), offset = log(n.f),
                   family = "poisson", data = df %>% filter(mate.type == "within" & trt.type == "expt")),
  
  all = glmer(n.coll ~ trt.minT*mate.ishs + (1|cage), offset = log(n.f),
              family = "poisson", data = df), # feels kinda wrong lol
  expt_btwn = glmer(n.coll ~ trt.minT * mate.ishs + (1|cage), offset = log(n.f),
                    family = "poisson", data = df %>% filter(mate.type == "between")),
  expt_all = glmer(n.coll ~ trt.minT * mate.ishs + (1|cage), offset = log(n.f),
                   family = "poisson", data = df %>% filter(trt.type == "expt"))
)

# trt effects
drop1(mods$all_win, test = "Chisq") # no diff btwn ctrl/hs
drop1(mods$expt_win, test = "Chisq") # no diff btwn hs groups

# sex effects
drop1(mods$all, test = "Chisq") # int is ns
drop1(update(mods$all, ~ . - trt.minT:mate.ishs), test = "Chisq") # ME are ns

#drop1(mods$expt_btwn, test = "Chisq") # int is ns
drop1(update(mods$expt_btwn, ~ . - trt.minT:mate.ishs), test = "Chisq") # sex is barely sig

#drop1(mods$expt_all, test = "Chisq") # interaction is ns
drop1(update(mods$expt_all, ~ . - trt.minT:mate.ishs), test = "Chisq") # no effect of minT or hs!
```


## eggs hatched

### main

```{r box and whisker}
pd <- 0.75

gg <- dfs_viz$eggs_n %>% #mutate(log = log(n.coll)) %>% View()
  ggplot(aes(y = p.hatch.perc)) +
  geom_boxplot(position = position_dodge(width = pd),
               outlier.shape = NA
               ) +
  geom_point(data = dfs_viz$eggs_n,
             aes(size = log(n.coll)),
             position = position_dodge(width = pd),
             alpha = 0.5) +
  scale_radius(range = c(1, 6)) +
  #scale_size_area(max_size = 6) +
  labs(y = "percent eggs hatched",
       size = "ln[total eggs collected]",
       ) +
  facet_wrap(~year) +
  theme_bw() +
  theme(legend.position = "top",
        legend.box = "vertical",
        legend.margin = margin()) +
  guides(color = guide_legend(order = 1),
         size = guide_legend(order = 2))

# x as minT
gg +
  aes(x = trt.minT, col = trt.mate,) +
  #geom_point(aes(group = trt.mate)) +
  scale_color_manual(values = p_scales$cols_trtsex, labels = p_scales$labs_trtmate) +
  scale_x_discrete(labels = p_scales$labs_trt) +
  labs(col = "heat-treated parent",
       x = p_scales$xlab_minT)
#ResToFig("fit-hatch-bytrt")

# x as trted sex
gg + 
  aes(col = trt.minT, x = trt.mate) +
  #geom_point(aes(group = trt.minT)) +
  scale_color_manual(values = p_scales$cols_minT, labels = p_scales$labs_minT) +
  scale_x_discrete(labels = p_scales$labs_trtmate) +
  labs(col = p_scales$xlab_minT,
       x = "heat-treated parent")
#ResToFig("fit-hatch-bysex")
```

### stats
```{r eggs hatched 2023}
# 2023
df <- dfs_stats$eggs %>%
  filter(year == 2023) %>%
  mutate(n.fail = n.coll - n.hatch) # otherwise breaks performance()

# for mods that threw singular with cage, switched to fixed (per oberpiller et al 2022 recommendations)
mods <- list(
  all_win = glmer(cbind(n.hatch, n.fail) ~ trt.type + (1|cage),
                  family = "binomial", data = df %>% filter(mate.type == "within")),
  # singular w cage, switch to fixed
  expt_win = glm(cbind(n.hatch, n.fail) ~ trt.minT,
                   family = "binomial", data = df %>% filter(mate.type == "within" & trt.type == "expt")), 
  
  # check_model() does not like these
  all = glm(cbind(n.hatch, n.fail) ~ trt.minT + mate.ishs, # still singular but bc of 426 having 0's
              family = "binomial", data = df),
  expt_btwn = glm(cbind(n.hatch, n.fail) ~ trt.minT,
                    family = "binomial", data = df %>% filter(mate.type == "between")),
  expt_all = glm(cbind(n.hatch, n.fail) ~ trt.minT * mate.ishs,
                   family = "binomial", data = df %>% filter(trt.type == "expt"))
)



# trt effects
drop1(mods$all_win, test = "Chisq") # no diff btwn ctrl/hs 
drop1(mods$expt_win, test = "Chisq") # tho minT is sig?
summary(mods$expt_win)

# sex effects
drop1(mods$all, test = "Chisq") # minT and sex are sig

drop1(mods$expt_btwn, test = "Chisq") # minT is sig

drop1(mods$expt_all, test = "Chisq") # interaction is sig
drop1(update(mods$expt_all, ~ . - trt.minT:mate.ishs), test = "Chisq") # and main effects
```

```{r eggs hatched 2025}
# 2025
df <- dfs_stats$eggs %>%
  filter(year == 2025) %>%
  mutate(n.fail = n.coll - n.hatch) # otherwise breaks performance()

mods <- list(
  all_win = glmer(cbind(n.hatch, n.fail) ~ trt.type + (1|cage),
                  family = "binomial", data = df %>% filter(mate.type == "within")),
  expt_win = glm(cbind(n.hatch, n.fail) ~ trt.minT,
                   family = "binomial", data = df %>% filter(mate.type == "within" & trt.type == "expt")), 
  
  all = glmer(cbind(n.hatch, n.fail) ~ trt.minT + mate.ishs + (1|cage),
              family = "binomial", data = df),
  # converges
  expt_btwn = glmer(cbind(n.hatch, n.fail) ~ trt.minT*mate.ishs + (1|cage),
                    family = "binomial", data = df %>% filter(mate.type == "between")),
  expt_btwn2 = glmer(cbind(n.hatch, n.fail) ~ mate.ishs + (1|cage),
                    family = "binomial", data = df %>% filter(mate.type == "between")),
  # converges
  expt_all = glmer(cbind(n.hatch, n.fail) ~ trt.minT * mate.ishs + (1|cage),
                   family = "binomial", data = df %>% filter(trt.type == "expt"))
)

# trt effects
drop1(mods$all_win, test = "Chisq") # no diff btwn ctrl/hs 
drop1(mods$expt_win, test = "Chisq") # tho minT is sig?

# sex effects
drop1(mods$all, test = "Chisq") # minT is sig

drop1(mods$expt_btwn, test = "Chisq") # int is ns
drop1(update(mods$expt_btwn, ~ . - trt.minT:mate.ishs), test = "Chisq") # ME are ns

drop1(mods$expt_all, test = "Chisq") # int is ns
drop1(update(mods$expt_all, ~ . - trt.minT:mate.ishs), test = "Chisq") # minT is sig
```




## sample size tables

### egg counts
```{r}
dfs_viz$eggs_ss %>%
  select(c(year, trt.minT, trt.mate, 
           starts_with(c("n.", "avg.", "se.")))) #%>% ResToCsv("egg results")
```

## [ ] sample sizes
```{r}
# TODO ideally use the same as the viz but maybe need to use the stats one lol cuz its on the per-tent level
# need to check the colony filtering on this...

# per cage
dfs_stats$eggs %>%
  group_by(year, trt.minT, trt.mate, cage) %>%
  summarise(perc.f = sum(n.f),
            perc.m = sum(n.m),
            perc.rat = n.f/n.m,
            perc.coll = sum(n.coll),
            perc.hatch = sum(n.hatch)) #%>% ResToCsv("tents/n/percage")

# per trt combo
dfs_stats$eggs %>%
  mutate(rat = n.f/n.m,
         tot.ads = n.f + n.m) %>%
  group_by(year,  trt.minT, trt.mate) %>%
  summarise(n = n(),
            tot.f = sum(n.f),
            tot.m = sum(n.m),
            avg.rat = mean(rat),
            avg.ads = mean(tot.ads),
            tot.coll = sum(n.coll),
            tot.hatch = sum(n.hatch)) %>% View() #%>% ResToCsv("tents/n/pertrt")
```

# ## dump (ref only)

## supp dev patchwork figs

post-mortem: will use variable/aesthetic pairings that match what's being done in the main figs instead!


```{r eval = FALSE}
# add facet var
dfs_viz[c("wide", "long")] <- dfs_viz[c("wide", "long")] %>%
  lapply(., \(x){
    x %>%
      mutate(trt.facet = case_when(trt == 260 ~ "ctrl",
                                   pop == "lab" ~ "lab",
                                   pop == "field" ~ "field"),
             trt.facet = factor(trt.facet))
  })

# legend fn for patchwork plots (trt.type/yr)
AddLeg_Patchwork <- function(){
  list(
    labs(lty = "population", shape = "population"),
    scale_shape_manual(values = p_scales$shp_pop),
    scale_linetype_manual(values = p_scales$lty_pop),
    guides(shape = guide_legend(override.aes = list(linetype = c("255F", "solid"))))
  )
}
```

```{r supp fig pup survival patchwork, eval = FALSE}
.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = prop.pup,
               shape = pop, lty = pop)) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = prop.pup + se.pup,
                      ymin = prop.pup - se.pup),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    facet_wrap(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "% survived to pupation") +
    AddLeg_Patchwork() +
    scale_color_brewer(palette = "Dark2") +
    theme_bw()
}

p1 <- dfs_viz$wide %>%
  filter(trt == 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = year, 
      group = pop)

p2 <- dfs_viz$wide %>%
  filter(trt != 260) %>%
  CalcSurvProps_def() %>%
  .plotfn() +
  aes(x = trt,
      group = year,
      col = year) +
  labs(x = p_scales$xlab_minT) +
  scale_x_discrete(labels = p_scales$labs_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0, 1))
```

```{r supp fig pup dev rate patchwork, eval = FALSE}
df <- dfs_viz$long %>%
  CalcDevSS_ad() %>%
  filter(instar == "pupa")

.plotfn <- function(x){
  x %>% 
    ggplot(aes(y = avg.devrate,
               shape = pop, lty = pop
               )) +
    geom_point(size = 2) +
    geom_line() +
    geom_errorbar(aes(ymax = avg.devrate + se.devrate,
                      ymin = avg.devrate - se.devrate),
                  width = 0.25, lty = "solid", show.legend = FALSE) +
    labs(y = "larval development rate (1/days)") +
    scale_color_brewer(palette = "Dark2") +
    AddLeg_Patchwork() +
    theme_bw()
}

p1 <- df %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year, group = pop)

p2 <- df %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt, group = interaction(pop, year), col = year) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  labs(x = p_scales$xlab_minT)

p1 + p2 +
  plot_layout(widths = c(1, 2),
              guides = "collect", axes = "collect") &
  ylim(c(0.029, 0.054))
```

```{r supp fig adult longevity patchwork, eval = FALSE}
df <- dfs_viz$long %>%
    filter(if.fridge == 0,
           instar == "long") %>%
    CalcDevSS_ad()

.plotfn <- function(x){
  x %>%
    ggplot(aes(y = avg.tt,
               shape = pop, lty = pop
               )
           ) +
    geom_errorbar(aes(ymin = avg.tt - se.tt,
                      ymax = avg.tt + se.tt),
                  width = 0.5, lty = "solid", show.legend = FALSE) +
    geom_point(size = 2) + 
    geom_line() +
    facet_grid(~ trt.facet, labeller = as_labeller(p_scales$facs_trtpop)) +
    labs(y = "adult longevity (days)",) +
    theme_bw() +
    AddLeg_Patchwork() +
    ylim(c(4, 17.5))
}

p1 <-  df %>%
  filter(trt == 260) %>%
  .plotfn() +
  aes(x = year,
      group = pop) +
  labs(x = "year")

p2 <- df %>%
  filter(trt != 260) %>%
  .plotfn() +
  aes(x = trt,
      col = year,
      group = interaction(year, pop)) +
  scale_x_discrete(labels = p_scales$labs_minT) +
  scale_color_brewer(palette = "Dark2") +
  labs(x = p_scales$xlab_minT)

p1 + p2 +
  plot_layout(widths = c(1,2),
              axes = "collect", guides = "collect")
```