---
title: "ntw mss posthoc"
date: "2026-08-09"
---

# preamble and setup

main script was getting way too long

- READING DUMP
  - https://rcompanion.org/rcompanion/d_08.html


```{r pathfinding, message=FALSE}
here::i_am("scripts/_mss/posthoc.Rmd")
```

- for NTW, models fit based on final models from `final-res.Rmd`. for CTRL, checked if int was sig.
- if int was sig in either NTW or CTRL models, did simple contrasts. if int was ns, used additive model and got EMMs per factor. 

## loading

```{r data and lib, message=FALSE}
library(emmeans) # post-hoc tests
source(here::here("scripts/_mss/load.R"))

# create df of just same minT group
dfs_stats <- list_modify(
  dfs_stats,
  dev_ph = dfs_stats$dev_all %>% filter(trt.minT == 26)
)
```

```{r}
# maybe make these objects for exporting instead lol...
# instead of %>% tidy %>% write_clip
```

  
# # survival

## pupal

```{r}
# NTW
mods <- glmer(is.pup ~ pop*trt.minT + (1|year),
             family = "binomial", data = dfs_stats$dev_expt)

anova(mods, update(mods, . ~ . - pop:trt.minT),
      test = "Chisq") # int is sig
drop1(update(mods, . ~ . - pop:trt.minT), test = "Chisq") # both are sig

emms <- emmeans(mods, c("pop", "trt.minT"), type = "response")

# pairs(emms, simple = "pop") 
# pairs(emms, simple = "trt.minT")
pairs(emms, simple = "each", combine = TRUE) 
```

```{r}
# CTRL VS EXPT
# fit
mods <- list(
  int = glmer(is.pup ~ pop*trt.type + (1|year),
             family = "binomial", data = dfs_stats$dev_ph),
  add = glmer(is.pup ~ pop + trt.type + (1|year),
             family = "binomial", data = dfs_stats$dev_ph)
)

# check main effs and int
anova(mods$add, mods$int, test = "Chisq") # int is sig
drop1(mods$add, test = "Chisq") # pop is sig

# get means and compare
emms <- emmeans(mods$int, c("pop", "trt.type"), type = "response")

# pairs(emms, simple = "pop") 
# pairs(emms, simple = "trt.type")
pairs(emms, simple = "each", combine = TRUE)
```

## adult

```{r}
# NTW
mods <- glmer(is.ec ~ pop*trt.minT + (1|year),
             family = "binomial", data = dfs_stats$dev_expt)

anova(mods, update(mods, . ~ . - pop:trt.minT),
      test = "Chisq") # int is ns
drop1(update(mods, . ~ . - pop:trt.minT), test = "Chisq") # pop is sig

# get emms separately
emms <- list(
  pop = emmeans(mods, "pop", type = "response"),
  trt = emmeans(mods, "trt.minT", type = "response")
  )

pairs(emms$pop)
pairs(emms$trt)
rbind(pairs(emms$pop), pairs(emms$trt))
```

```{r}
# CTRL VS EXPT
# fit
mods <- list(
  int = glmer(is.ec ~ pop*trt.type + (1|year),
             family = "binomial", data = dfs_stats$dev_ph),
  add = glmer(is.ec ~ pop + trt.type + (1|year),
             family = "binomial", data = dfs_stats$dev_ph)
)
## both singular fits - not super surprising bc lab generally has high survival

# check main effs and int
anova(mods$add, mods$int, test = "Chisq") # int is barely sig
drop1(mods$add, test = "Chisq") # pop is sig
## int is the better model, but bc int is borderline, get EMMs on the int mod for 1 factor at a time

# get means and compare
emms <- list(
  pop = emmeans(mods$int, "pop", type = "response"),
  trt = emmeans(mods$int, "trt.type", type = "response")
)
  
pairs(emms$pop) 
pairs(emms$trt)
rbind(pairs(emms$pop), pairs(emms$trt))
```


# # mass

## pupal

```{r}
# NTW
mods <- lmer(mass.pupa ~ pop*trt.minT + sex + (1|year),
             data = dfs_stats$dev_expt %>% filter(!is.na(sex)))

anova(mods, type = 2) # all sig

emms <- list(
  int = emmeans(mods, c("pop", "trt.minT")),
  sex = emmeans(mods, "sex")
  )
emms$int


pairs(emms$int, simple = "pop")
pairs(emms$int, simple = "trt.minT")
pairs(emms$int, simple = "each", combine = TRUE)
pairs(emms$sex)

rbind(pairs(emms$int, simple = "each", combine = TRUE), pairs(emms$sex))

# visualise...
with_emm_options(gg.theme = ggplot2::theme_bw(),
                 emmip(mods, pop ~ trt.minT | sex, 
      CIs = TRUE, 
      ))

with_emm_options(gg.theme = ggplot2::theme_bw(),
                 emmip(mods, trt.minT ~ pop | sex, 
      CIs = TRUE, 
      ))
```

```{r}
# CTRL VS EXPT
# fit
mods <- lmer(mass.pupa ~ pop*trt.type + sex + (1|year),
             data = dfs_stats$dev_ph %>% filter(!is.na(sex)))

# check main effs and int
anova(mods, type = 2) # all are sig

# get means and compare
emms <- list(
  int = emmeans(mods, c("pop", "trt.type")),
  sex = emmeans(mods, "sex")
  )

# pairs(emms$int, simple = "pop")
# pairs(emms$int, simple = "trt.type")
# pairs(emms$sex)
rbind(pairs(emms$int, simple = "each", combine = TRUE), pairs(emms$sex))
```

## adult

```{r}
# NTW
mods <- lmer(mass.eclose ~ pop*trt.minT + sex + (1|year),
             data = dfs_stats$dev_expt %>% filter(!is.na(sex)))
anova(mods, type = 2) # all sig

emms <- list(
  int = emmeans(mods, c("pop", "trt.minT")),
  sex = emmeans(mods, "sex")
  )

pairs(emms$int, simple = "pop")
pairs(emms$int, simple = "trt.minT")
pairs(emms$sex)

rbind(pairs(emms$int, simple = "each", combine = TRUE), pairs(emms$sex))
```

```{r}
# CTRL VS EXPT
# fit
mods <- lmer(mass.eclose ~ pop*trt.type + sex + (1|year),
             data = dfs_stats$dev_ph %>% filter(!is.na(sex)))

# check main effs and int
anova(mods$int, type = 2) # all are sig

# get means and compare
emms <- list(
  int = emmeans(mods, c("pop", "trt.type")),
  sex = emmeans(mods, "sex")
  )

pairs(emms$int, simple = "pop")
pairs(emms$int, simple = "trt.type")
pairs(emms$sex)

rbind(pairs(emms$int, simple = "each", combine = TRUE), pairs(emms$sex))
```


# # dev time

## pupal

```{r}
# NTW
mods <- lmer(tt.pupa ~ pop*trt.minT + (1|year),
             data = dfs_stats$dev_expt)
anova(mods, type = 2) # ME are sig, int is ns

# get EMMs
emms <- list(
  pop = emmeans(mods, "pop"),
  trt = emmeans(mods, "trt.minT")
)
  
pairs(emms$pop)
pairs(emms$trt)

rbind(pairs(emms$pop), pairs(emms$trt))
```

```{r}
# CTRL VS EXPT
# fit
mods <- lmer(tt.pupa ~ pop*trt.type + (1|year),
             data = dfs_stats$dev_ph)

# check main effs and int
anova(mods, type = 2) # all sig

# get means and compare
emms <- emmeans(mods, c("pop", "trt.type"))

# pairs(emms, simple = "pop")
# pairs(emms, simple = "trt.type")

pairs(emms, simple = "each", combine = TRUE)
```

## adult

```{r}
# NTW
mods <- lmer(tt.eclose ~ pop*trt.minT + (1|year),
             data = dfs_stats$dev_expt)
anova(mods, type = 2) # trt is sig, int is ns

# get EMMs
emms <- list(
  pop = emmeans(mods, "pop"),
  trt = emmeans(mods, "trt.minT")
)
  
# pairs(emms$pop) %>% tidy()
# pairs(emms$trt) %>% tidy()

rbind(pairs(emms$pop), pairs(emms$trt))
```

```{r}
# CTRL VS EXPT
# fit
mods <- lmer(tt.eclose ~ pop*trt.type + (1|year),
             data = dfs_stats$dev_ph)

# check main effs and int
anova(mods, type = 2) # all sig

# get means and compare
emms <- emmeans(mods, c("pop", "trt.type"))

# pairs(emms, simple = "pop")
# pairs(emms, simple = "trt.type")

pairs(emms, simple = "each", combine = TRUE)
```

# # lifespan

## 2023
```{r}
# NTW
mods <- lm(tt.long ~ pop*trt.minT,
           data = dfs_stats$dev_expt %>% filter(year == 2023, 
                                                if.mated == 1, if.fridge == 0))

drop1(mods, test = "F") # int is ns, so use the additive
mods <- update(mods, ~ . - pop:trt.minT)

drop1(mods, test = "F") # everything ns

emms <- list(
  pop = emmeans(mods, "pop"),
  trt = emmeans(mods, "trt.minT")
  )

pairs(emms$pop)
pairs(emms$trt)
rbind(pairs(emms$pop), pairs(emms$trt))
```

```{r}
# CTRL
mods <- lm(tt.long ~ pop*trt.type,
           data = dfs_stats$dev_ph %>% filter(year == 2023, 
                                              if.mated == 1, if.fridge == 0))

drop1(mods, test = "F") # int is ns

mods <- update(mods, ~ . - pop:trt.type)
drop1(mods, test = "F") # pop is sig

emms <- list(
  pop = emmeans(mods, "pop"),
  trt = emmeans(mods, "trt.type")
  )
test(emms$pop, adjust = "bonferroni") %>% tidy() %>% write_clip()
test(emms$trt, adjust = "bonferroni") %>% tidy() %>% write_clip()

pairs(emms$pop)
pairs(emms$trt)
rbind(pairs(emms$pop), pairs(emms$trt)) 
```

## 2025

```{r}
# fridge effects lol
mods <- lm(tt.long ~ pop*trt.minT + if.fridge, 
             data = dfs_stats$dev_expt %>% filter(year == 2025, if.mated == 1))

drop1(mods, test = "F")

emms <- emmeans(mods, "if.fridge")
pairs(emms)
```


```{r}
# NTW
mods <- lm(tt.long ~ pop*trt.minT,
           data = dfs_stats$dev_expt %>% filter(year == 2025, 
                                                if.mated == 1, if.fridge == 0))

drop1(mods, test = "F") # int is ns

mods <- update(mods, ~ . - pop:trt.minT)
drop1(mods, test = "F") # trt is sig

emms <- list(
  pop = emmeans(mods, "pop"),
  trt = emmeans(mods, "trt.minT")
  )

pairs(emms$pop)
pairs(emms$trt)
rbind(pairs(emms$pop), pairs(emms$trt))
```

```{r}
# CTRL
mods <- lm(tt.long ~ pop*trt.type,
           data = dfs_stats$dev_ph %>% filter(year == 2025, 
                                              if.mated == 1, if.fridge == 0))

drop1(mods, test = "F") # int is ns so use additive

mods <- update(mods, ~ . - pop:trt.type)
drop1(mods, test = "F") # all ns

emms <- list(
  pop = emmeans(mods, "pop"),
  trt = emmeans(mods, "trt.type")
  )

pairs(emms$pop)
pairs(emms$trt)
rbind(pairs(emms$pop), pairs(emms$trt))
```

# eggs laid
