---
title: "ntw 2025 stats analyses"
date: "2025-12-02"
---

# # preamble & loading

stats EDAs for ntw 2025.

organised by life stage/main metric (can sort of just keep appending down a section)

```{r set pathfinding}
here::i_am("scripts/analysis/stats-dev.Rmd")
```

```{r load libraries, message=FALSE}
library(here)
library(lme4) # glm, lm, glmer
#library(lmerTest) # lmer
#library(MuMIn) # dredge (nested model comparisons)
library(tidyverse)

library(conflicted)
conflicts_prefer( 
  #lmerTest::lmer, # dont rlly need..
  dplyr::filter
)
# lmerTest::lmer overloads lme4::lmerMod summary() and anova() to provide p-vals for the fixed effs

# library(survival) # for fitting 
# library(survminer) # for plotting
# library(coxme) # mixed effects coxphs (but also frailty() might be enough...)

source(here("scripts/analysis/utils/filters.R"))
source(here("scripts/analysis/utils/stats.R"))
```

notes
- `lmerTest` docs: https://www.jstatsoft.org/article/view/v082i13

```{r load data, warning=FALSE}
source(here("scripts/analysis/load-dev.R"))
```

## TODO

## df prep

notes
- on handling subsets:
  - for exptal, pred vars include `trt.minT`
  - for ctrl + exptal, pred vars include `trt.type` (i.e. HS or not)

```{r dev data prep}
# prep exptal only + all bugs subsets. 
# further subset/mutate as needed closer to the codeblock (ie mutates, more filtering, etc etc)

dfs_imported <- list(
  all = dfs_tidy$wide,
  expt = dfs_tidy$wide %>% filter(trt != 260)
) %>%
  lapply(., \(x) {
    x %>%
      filter(!(pop == "lab" & diet == "TB"),
             pop != "col",
             !is.na(is.pup), # drop culled bugs here
             ) %>%
      mutate(is.pup = case_when(is.pup == 2 ~ 1,
                                TRUE ~ is.pup), # treat LPI as pup
             across(c("year", "trt.minT", "trt.type", "trt"), as.factor),
             pop = factor(pop, levels = c("lab", "field")))
  })
```

# ### pupal 

## main results

### survival to pup

```{r pup surv model generation}
mods_pupsurv <- list(
  # set na.action here so can dredge later
  expt_fullint = glm(is.pup ~ year*pop*trt.minT, 
                     family = "binomial", data = dfs_imported$expt, 
                     na.action = na.fail),
  all_fullint = glm(is.pup ~ year*pop*trt.type,
                    family = "binomial", data = dfs_imported$all,
                    na.action = na.fail)
  )
```

- exptal: 1 and 2-way ints are signif

```{r pup surv model results}
mods_pupsurv %>%
  lapply(., GetModelResults)

# dredging: print out is long asf w/o subsetting
.modfn <- function(x){
  x %>% dredge() %>% subset(., weight > 0.05, recalc.weights = FALSE)
}

mods_pupsurv %>% lapply(., .modfn)
```

```{r write pup surv results, eval=FALSE}
# save anova + dredge separately
mods_pupsurv %>%
  lapply(., \(x){
    x %>%
      .modfn() %>%
      as.data.frame()
  }) %>%
  #reduce(full_join) %>% View() # same result as rbindlist bc of diff cols
  rbindlist(fill = TRUE, idcol = "subset") %>%
  WriteModelResultsCsv("pup-surv_dredge")

# update models; dropping unsig terms per initial anova (maybe ?)
#mods_pupsurv %>% # for pup-surv_anova.csv
list(
  expt_final = glm(is.pup ~ (year + pop+ trt.minT)^2, 
                     family = "binomial", data = dfs_imported$expt),
  all_final = mods_pupsurv$all_fullint
) %>%
  lapply(., \(x){
    x %>% 
      anova(test = "Chisq") %>%
      as.data.frame() %>%
      mutate(vars = rownames(.), .before = 1)
  }) %>%
  rbindlist(fill = TRUE, idcol = "subset") %>% #View()
  WriteModelResultsCsv("pup-surv_anova-final")
```


### pupal mass

```{r}
dfs <- dfs_imported %>%
  lapply(., \(x) {
    x %>% filter(!is.na(mass.pupa),
                 !is.na(sex))
  })
```

- exptal: 1-ways, `year:minT`, and `minT:pop` are sig

```{r pup mass model generation}
# set na.action here to dredge later

mods_pupmass <- list(
  expt_fullint = lm(mass.pupa ~ year*trt.minT*pop*sex,
               data = dfs$expt, na.action = na.fail),
    # subsets of the full model are in a later chunk otherwise results r crazy
  all_fullint = lm(mass.pupa ~ year*trt.type*pop*sex,
                    data = dfs$all, na.action = na.fail)
  )
```

```{r pup mass model main results}
mods_pupmass %>%
  lapply(., GetModelResults)

# dredging. output is long w/o the subsetting
# (same subsetting as pup survival)
# .modfn <- function(x){
#   x %>% dredge() %>% subset(., weight > 0.05, recalc.weights = FALSE)
# }

mods_pupmass %>% lapply(.modfn)
```

```{r write pup mass results, eval=FALSE}
# (write csv before doing other additional anovas bc mods df gets overwritten)

mods_pupmass %>%
  lapply(., \(x){
    x %>% 
      anova(test = "Chisq") %>%
      as.data.frame() %>%
      mutate(vars = rownames(.), .before = 1)
  }) %>%
  rbindlist(fill = TRUE, idcol = "subset") %>%
  WriteModelResultsCsv("pup-mass_anova")

mods_pupmass %>% %>% lapply(., \(x){
    x %>%
      .modfn() %>%
      as.data.frame()
    }) %>%
  rbindlist(fill = TRUE, idcol = "subset") %>%
  WriteModelResultsCsv("pup-mass_dredge")
```

(more commentary on results)

*exptal:*
basically, per anovas, `lm(mass.pupa ~ year + minT + pop + sex + year:minT)` seems good...

other results: 

- bc the dredge of the full interactive model (`year*pop*minT*sex`) is crazy, how do the subsets compare?
  - doing `anova(simpler, complex)` per https://www.geeksforgeeks.org/machine-learning/comparing-two-linear-models-with-anova-in-r/ 

```{r expt pup mass model subsets, eval=FALSE}
# keep separate otherwise console blows up

mods_pupmass <- list_modify(
  mods_pupmass,
  expt_add = lm(mass.pupa ~ year + trt.minT + pop + sex,
                data = dfs$expt, na.action = na.fail),
  # subsets of the full model
  expt_2way = lm(mass.pupa ~ (year + trt.minT + pop + sex)^2,
              data = dfs$expt, na.action = na.fail),
  expt_3way = lm(mass.pupa ~ year*trt.minT*pop + sex,
              data = dfs$expt, na.action = na.fail)
)

# testing anova(simplermod, complexmod)
mods_pupmass$expt_add %>% GetModelResults()
anova(mods_pupmass$expt_add, mods_pupmass$expt_fullint, test = "Chisq")
anova(mods_pupmass$expt_2way, mods_pupmass$expt_fullint, test = "Chisq")
anova(mods_pupmass$expt_3way, mods_pupmass$expt_fullint, test = "Chisq")
anova(mods_pupmass$expt_2way, mods_pupmass$expt_3way, test = "Chisq")
```


### time to pup

```{r pup time df prep}
dfs <- dfs_imported %>%
  lapply(., \(x) {
    x %>% 
      #filter(!is.na(jdate.pupa)) %>%
      mutate(#tt.pupa = jdate.pupa - jdate.hatch,
             devrate.pupa = 1/tt.pupa) %>%
      filter(!is.na(tt.pupa))
  })
```


```{r pup time model generation}
mods_puptime <- list(
  expt_fullint = lm(devrate.pupa ~ year*trt.minT*pop, 
                    data = dfs$expt, na.action = na.fail),
  all_fullint = lm(devrate.pupa ~ year*trt.type*pop, 
                   data = dfs$all, na.action = na.fail)
  )
```

- 1 ways and `year:minT`, `year:pop` are sig

notes
- checked that `sex` PROBABLY doesnt matter: bc sex is only determined after pupation anyway... (so omit from mods)

```{r pup time model main results}
mods_puptime %>%
  lapply(., GetModelResults)

# use same dredge fn as survival
# .modfn <- function(x){
#   x %>% dredge() %>% subset(., weight > 0.05, recalc.weights = FALSE)
# }

mods_puptime %>% lapply(., .modfn)
```

```{r write pup time results, eval=FALSE}
mods_puptime %>%
  lapply(., \(x){
    x %>%
      anova(test = "Chisq") %>%
      as.data.frame() %>%
      mutate(vars = rownames(.), .before = 1)
  }) %>%
  rbindlist(fill = TRUE, idcol = "subset") %>%
  WriteModelResultsCsv("pup-time_anova")

mods_puptime %>%
  lapply(., \(x){
    x %>%
      .modfn() %>%
      as.data.frame()
  }) %>%
  rbindlist(fill = TRUE, idcol = "subset") %>%
  WriteModelResultsCsv("pup-time_dredge")
  
```

## EDA: other terms
2026-07-28 what terms to include?

### dev time

```{r}
mods <- list(
  m1 = lm(tt.pupa ~ pop*trt.minT*year, data = dfs_imported$expt),
  m2 = lm(tt.pupa ~ pop*trt.minT*year + as.factor(is.sup), data = dfs_imported$expt)
)

anova(mods$m1, mods$m2)
AIC(mods$m1, mods$m2) #including sup is better

summary(mods$m1)
summary(mods$m2)
```
### mass
```{r}
df <- dfs_imported$expt %>%
  filter(!is.na(sex))

mods <- list(
  m1 = lm(mass.pupa ~ pop*trt.minT*year, data = df),
  m2 = lm(mass.pupa ~ pop*trt.minT*year*sex, data = df),
  m3 = lm(mass.pupa ~ pop*trt.minT*year + sex, data = df)
)

anova(mods$m1, mods$m3, mods$m2) # adding sex (1way) is better
AIC(mods$m1, mods$m2, mods$m3)
summary(mods$m1)
summary(mods$m3)
```



## EDA: random yr (2)
2026-07-27 trying this again after doing some more reading...

compare if models with yr as random are better. (just use the full interaction model bc easier to compare w the null this way)

### pup surv

```{r}
mods <- list(
  m0 = glm(is.pup ~ pop*trt.minT, family = "binomial", data = dfs_imported$expt),
  fm = glm(is.pup ~ pop*trt.minT*year, family = "binomial", data = dfs_imported$expt),
  mm = glmer(is.pup ~ pop*trt.minT + (1|year), family = "binomial", data = dfs_imported$expt)
)

anova(mods$m0, mods$fm, test = "Chisq") # include year
anova(mods$mm, mods$m0) # include year 

anova(mods$mm, mods$fm, mods$m0)
anova(mods$mm, mods$fm)
summary(mods$mm)
summary(mods$fm)
1 - (1155.6/1457.1) # pseudo R2 = 0.2069179
1 - pchisq(1155.6, 1144) # p-val = 0.399

# per AIC, use the fixed
```

### dev time
```{r}
mods <- list(
  m0 = lm(tt.pupa ~ pop*trt.minT, data = dfs_imported$expt),
  fm = lm(tt.pupa ~ pop*trt.minT*year, data = dfs_imported$expt),
  mm = lmer(tt.pupa ~ pop*trt.minT + (1|year), data = dfs_imported$expt)
)

anova(mods$mm, mods$fm, mods$m0)

summary(mods$mm)
summary(mods$fm)
```


### pup mass

```{r}
mods <- list(
  m0 = lm(mass.pupa ~ pop*trt.minT + sex, data = dfs_imported$expt),
  fm = lm(mass.pupa ~ pop*trt.minT*year + sex, data = dfs_imported$expt),
  mm = lmer(mass.pupa ~ pop*trt.minT + sex + (1|year), data = dfs_imported$expt)
)

anova(mods$mm, mods$fm, mods$m0)

summary(mods$mm)
summary(mods$fm)

# fixed is barely better than mixed!
```





## EDA: random yr (1)
(for exptal bugs only) looking at effect of `year` as a random effect on response variables

per 251217 wkly mtg, keep year as a fixed effect in the models (but here is some EDA for reference just in case.)

each chunk is self-contained (model fitting + tests) and overwrites the `mods_pup_randyr` object. just run `GetModelResults` on them to be reminded of what happened)

### survival
```{r pup surv rand yr, eval=FALSE}
### model generation (just focus on expt bugs)
dfs <-  dfs_imported$expt %>%
  mutate(year_n = as.numeric(year))

mods_pup_randyr <- list(
  surv = glmer(is.pup ~ pop*trt.minT + (1|year),
                     family = "binomial", data = dfs),
  
  surv_ctsyr = glmer(is.pup ~ pop*trt.minT + (1|year_n),
                     family = "binomial", data = dfs),
    # compare to `surv` to examine effs of yr as cts or cat
  
  surv_noyr = glm(is.pup ~ pop*trt.minT,
                     family = "binomial", data = dfs)
    # compare to `surv` to examine effs of year
)


### tests
# effect of cts vs cat yr
anova(mods_pup_randyr$surv, mods_pup_randyr$surv_ctsyr, test = "Chisq") 

# to get approx p-value for (1|year)?:
anova(mods_pup_randyr$surv, mods_pup_randyr$surv_noyr, test = "Chisq") 
  # i forget what this means...
```

notes about `(1|year)` on pup surv:

- `glmer(is.pup ~ pop*minT + (1|year))` results are same if `year` is numeric or factor
- can approximate the p-value of the random effect by using `anova` to compare models with and without the effect of interest (bc they're nested) -- see more https://stats.stackexchange.com/questions/381208/r-how-to-get-estimates-and-p-values-for-random-effects-in-glmer

### dev time
```{r pup dev time rand yr, eval=FALSE}
dfs <- dfs_imported %>%
  lapply(., \(x) {
    x %>% 
      filter(!is.na(jdate.pupa)) %>%
      mutate(#tt.pupa = jdate.pupa - jdate.hatch,
             devrate.pupa = 1/tt.pupa)
  })

# mostly reminding myself of how lmerTest works
mods_pup_randyr <- list(
  time_lmertest = lmerTest::lmer(tt.pupa ~ trt.minT*pop + (1|year), data = dfs$expt),
  time_lme4 = lme4::lmer(tt.pupa ~ trt.minT*pop + (1|year), data = dfs$expt)
)

mods_pup_randyr %>% lapply(., summary)
```

### mass
```{r pup mass rand yr, eval=FALSE}
dfs <- dfs_imported %>%
  lapply(., \(x) {
    x %>% 
      filter(!is.na(jdate.pupa),
             !is.na(sex),
             !is.na(mass.pupa)
      )
    })

mods_pup_randyr <- list(
  mass = lmer(mass.pupa ~ trt.minT*pop*sex + (1|year),
              data = dfs$expt)
)
```


# ### adult

## EDA: other terms
2026-07-28 what other terms to include?

### surv to ec
```{r}
df <- dfs_imported$expt %>%
  filter(!is.na(mass.pupa))

mods <- list(
  m1 = glm(is.ec ~ year*pop*trt.minT,
           family = "binomial", data = df),
  m2 = glm(is.ec ~ year*pop*trt.minT + mass.pupa,
           family = "binomial", data = df)
)

anova(mods$m1, mods$m2, test = "Chisq") # better to add pupal mass
AIC(mods$m1, mods$m2)

summary(mods$m1)
summary(mods$m2)
```

### eclosion mass

```{r}
df <- dfs_imported$expt %>%
  filter(!is.na(mass.pupa))

# incl sex, pupal mass?
mods <- list(
    # w/ sex
    m1 = lm(mass.eclose ~ year*pop*trt.minT*sex,
           data = df),
    m1a = lm(mass.eclose ~ year*pop*trt.minT*sex + mass.pupa,
            data = df),
    # w/o sex
    m2 = lm(mass.eclose ~ year*pop*trt.minT,
            data = df),
    m2a = lm(mass.eclose ~ year*pop*trt.minT + mass.pupa,
            data = df)
)

anova(mods$m2, mods$m2a, mods$m1, mods$m1a) 
  # adding sex is w/e but mass is impt
AIC(mods$m2, mods$m2a, mods$m1, mods$m1a)
  # adding mass reduces AIC a lot; technically so does sex
  # best is to have sex and mass pupa..
summary(mods$m2)
summary(mods$m1)
summary(mods$m1a)

# try sex as a random eff
mods <- list(
  m0 = lm(mass.eclose ~ year*pop*trt.minT + mass.pupa,
            data = df),
  m1 = lm(mass.eclose ~ year*pop*trt.minT*sex + mass.pupa,
            data = df),
  m2 = lmer(mass.eclose ~ year*pop*trt.minT + (1|sex) + mass.pupa,
            data = df) # warning: rank deficient
)

anova(mods$m2, mods$m1, mods$m0) # fixed is better...?
```



## EDA: random year

2026-07-28 like pup responses, compare if models with yr as random are better. (just use the full interaction model bc easier to compare w the null this way)

### survival to ec
```{r}
mods <- list(
  fm = glm(is.ec ~ year*pop*trt.minT,
           family = "binomial", data = dfs_imported$expt),
  m0 = glm(is.ec ~ pop*trt.minT,
           family = "binomial", data = dfs_imported$expt),
  mm = glmer(is.ec ~ pop*trt.minT + (1|year),
           family = "binomial", data = dfs_imported$expt)
) 

# fit is singular for glmer, sort of an issue bc LME assumes missing data is 
# "missing at random" (https://stats.stackexchange.com/questions/354279/can-a-linear-mixed-model-handle-missing-data-that-is-not-missing-at-random)

anova(mods$mm, mods$fm, mods$m0) # not diff from e/o tho

```

### time to ec

```{r}
mods <- list(
  m0 = lm(tt.eclose ~ pop*trt.minT, data = dfs_imported$expt),
  fm = lm(tt.eclose ~ pop*trt.minT*year, data = dfs_imported$expt),
  mm = lmer(tt.eclose ~ pop*trt.minT + (1|year), data = dfs_imported$expt)
)

anova(mods$mm, mods$fm, mods$m0) # need to include yr; fixed is better

summary(mods$fm)
```

### ec mass
```{r}
mods <- list(
  m0 = lm(mass.eclose ~ pop*trt.minT + sex, data = dfs_imported$expt),
  fm = lm(mass.eclose ~ pop*trt.minT*year + sex, data = dfs_imported$expt),
  mm = lmer(mass.eclose ~ pop*trt.minT + sex + (1|year), data = dfs_imported$expt)
)

anova(mods$mm, mods$fm, mods$m0) # need to include yr; fixed is better

summary(mods$fm)
```

### longevity
```{r}
df <- dfs_imported$expt %>%
  filter(if.mated == 1)
mods <- list(
  m0 = lm(tt.long ~ pop*trt.minT + sex, data = dfs_imported$expt),
  fm = lm(tt.long ~ pop*trt.minT*year + sex, data = dfs_imported$expt),
  mm = lmer(tt.long ~ pop*trt.minT + sex + (1|year), data = dfs_imported$expt)
)

anova(mods$mm, mods$fm, mods$m0) # need to include yr; fixed is better

summary(mods$fm)
```


## EDA: pupal mass

looking at effect of pupal mass on response variables. compare to models from final mss results.

### survival

```{r}
# on surv to ec
df <- dfs_imported$expt %>%
  filter(!is.na(mass.pupa))

mods <- list(
  m0 = glm(is.ec ~ year*pop*trt.minT + mass.pupa,
           family = "binomial", data = df),
  m0a = glm(is.ec ~ year*pop*trt.minT,
            family = "binomial", data = df),
  m1 = glm(is.ec ~ year*pop + mass.pupa, 
           family = "binomial", data = df),
  m2 = glm(is.ec ~ mass.pupa,
           family = "binomial", data = df)
)

AIC(mods$m0, mods$m0a, mods$m1, mods$m2)
anova(mods$m0, mods$m0a, mods$m1, mods$m2)

anova(mods$m0, test = "Chisq") # pop **, mass ***, pop*T **, 3way .
anova(mods$m0a, test = "Chisq") # pop **, year:pop .
anova(mods$m1, test = "Chisq")

# probably makes more sense to drop mass.. (anova agrees with fig...)
  ## and again, issues with unequal groups
```

### eclosion mass

```{r}
# on eclosion mass
df <- dfs_imported$expt %>%
  filter(!is.na(mass.pupa))

mods <- list(
  m0 = lm(mass.eclose ~ year*pop*trt.minT*sex + mass.pupa,
          data = df),
  m0a = lm(mass.eclose ~ year*pop*trt.minT + mass.pupa,
          data = df),
  m1 = lm(mass.eclose ~ year*pop*trt.minT*sex,
          data = df),
  m1a = lm(mass.eclose ~ year*pop*trt.minT,
          data = df)
)

anova(mods$m0, mods$m0a, mods$m1, mods$m1a) # models w sex are better?
AIC(mods$m0, mods$m0a, mods$m1, mods$m1a) # agree w/ AIC... (and w mass)
# so continue analyses w sex

anova(mods$m0)
anova(mods$m1)
# to simplify, can just look at 2 ways for both

mods <- list_modify(
  mods,
  m2 = lm(mass.eclose ~ (year + pop + trt.minT + sex)^2 + mass.pupa,
          data = df),
  m2a = lm(mass.eclose ~ (year + pop + trt.minT + sex)^2,
          data = df)
)

anova(mods$m2, mods$m2a) 
AIC(mods$m2, mods$m2a) # better to keep pupal mass in..?

anova(mods$m2) # 1 ways ***, year:sex ., trt:sex *
anova(mods$m2a) # year, pop, sex ***; trt, pop:trt, trt:sex **, pop:sex *
  # (ie 2 ways with year fall out)

# per figs, sex stuff is prob bc of unequal groups.
# again, feels somewhat better with removing pup mass?
```


## EDAs: longevity

looking at predictor variables on longevity

### effect of sex
```{r}
# include sex? parse the mating? drop 2024? (same qns as the viz)

mods <- list(
  m1 = lm(tt.long ~ year*pop*trt.minT*sex,
          data = dfs_imported$expt),
  m2 = lm(tt.long ~ year*pop*trt.minT*sex,
          data = dfs_imported$expt %>% filter(if.mated == 1))
)

anova(mods$m1) # can drop if looking at everyone
anova(mods$m2) # sex matters if looking at mated bugs only...
```

### effect of fridge 
```{r}
# per mss, drop 2024
dfs <- list(
  all = dfs_imported$all %>% filter(year != 2024),
  expt = dfs_imported$expt %>% filter(year != 2024)
)

mods <- list(
  m1 = lm(tt.long ~ year*pop*trt*as.factor(if.fridge),
          data = dfs$all), 
  m1a = lm(tt.long ~ year*pop*trt*as.factor(if.fridge),
          data = dfs$all %>% filter(if.mated == 1)),
  m2 = lm(tt.long ~ year*pop*trt.minT*as.factor(if.fridge),
          data = dfs$expt),
  m2a = lm(tt.long ~ year*pop*trt*as.factor(if.fridge),
          data = dfs$expt %>% filter(if.mated == 1))
)

anova(mods$m1) # all trts
anova(mods$m1a) # all trts, mated only
anova(mods$m2) # expt only
anova(mods$m2a) # expt, mated only

# fridge effects only fall out if looking at the mated, exptal bugs... (i.e. so can include them back in the results i guess?)

# for each year?
df <- dfs$expt %>%
  filter(if.mated == 1,
         !is.na(if.fridge))

mods <- list(
  y23a = lm(tt.long ~ pop*trt,
          data = df %>% filter(year == 2023)),
  y23b = lm(tt.long ~ pop*trt*as.factor(if.fridge),
          data = df %>% filter(year == 2023)),
  y25a = lm(tt.long ~ pop*trt,
          data = df %>% filter(year == 2025)),
  y25b = lm(tt.long ~ pop*trt*as.factor(if.fridge),
          data = df %>% filter(year == 2025))
)

anova(mods$y23a, mods$y23b) # doesnt matter in 2023
anova(mods$y25a, mods$y25b) # matters in 2025 (but prob cuz i kept better track..)

# with all data
mods <- list(
  y23 = lm(tt.long ~ pop*trt*as.factor(if.fridge),
          data = dfs$expt %>% filter(if.mated == 1, year == 2023)),
  y25 = lm(tt.long ~ pop*trt*as.factor(if.fridge),
          data = dfs$expt %>% filter(if.mated == 1, year == 2025))
)
anova(mods$y23)
anova(mods$y25) # fridge matters in 2025 though probably bc n=1 for some combinations
```

### fridge and sex?
```{r}
# per mss, drop 2024
dfs <- list(
  all = dfs_imported$all %>% filter(year != 2024),
  expt = dfs_imported$expt %>% filter(year != 2024)
)

mods <- list(
  m1 = lm(tt.long ~ year*sex*pop*trt*as.factor(if.fridge),
          data = dfs$expt),
  m1a = lm(tt.long ~ year*pop*trt + sex*as.factor(if.fridge),
           data = dfs$expt))

anova(mods$m1)
anova(mods$m1a)

# probably not that much.. (again, issues due to N=1)
```



# ### misc

## sample sizes
```{r}
# calc to pups by sex separately cuz needs pivots
list(
  dfs_imported$all %>%
    mutate(year = as.character(year)) %>%
    group_by(year, trt, pop) %>%
    summarise(n = n(),
              to.pup = sum(is.pup == 1)),
  
  dfs_imported$all %>%
    group_by(trt, pop) %>%
    summarise(n = n(),
           to.pup = sum(is.pup == 1),
           year = "all")
  ) %>%
  lapply(., \(x){
    x %>%
      pivot_wider(names_from = "trt", values_from = c("n", "to.pup")) %>%
      pivot_longer(cols = starts_with(c("n", "to.pup")),
                   names_to = c("count", ".value"),
                   names_sep = "_")
  }) %>%
  reduce(full_join)
```


## EDA: age vs instar

looking at dev trajectories: is `age ~ instar` linear?

```{r}
# df prep
dfs <- dfs_tidy$long %>%
  # pick out from 3rd-wander bc all have same trts
  filter(instar %in% c("3rd", "4th", "5th", 
                       "6th", "7th", "8th", 
                       "wander")) %>%
  mutate(across(c("year", "sup", starts_with(c("trt", "is."))),
                as.factor))
```

```{r}
# non-sups only
list(
  nosups = lm(tt ~ instar * trt, data = filter(dfs, is.sup == 0)),
  nosups2 = lm(tt ~ instar * trt * pop, data = filter(dfs, is.sup == 0))
  ) %>%
  GetModelResults()







```

