---
title: "ntw 2025 stats analyses"
date: "2025-12-02"
---

# # preamble & loading

stats analyses for ntw 2025

```{r load libraries and utils}
here::i_am("scripts/analysis/stats-tents.Rmd")
library(here)

library(lme4) # glmer, glmer.nb (mixed effs)
library(lmerTest) # lmer
## lmerTest::lmer overloads lme4::lmerMod summary() and anova() to provide p-vals for the fixed effs
#library(MuMIn) # dredge (nested model comparisons)
library(car) # Anova
library(pscl) # zeroinfl
library(MASS) # glm.nb

library(tidyverse)

select <- dplyr::select
  ## if MASS:select is misbehaving

library(conflicted)
conflicts_prefer(
  lmerTest::lmer,
  dplyr::filter(),
  dplyr::select()
)

# library(survival) # for fitting 
# library(survminer) # for plotting
# library(coxme) # mixed effects coxphs (but also frailty() might be enough...)

source(here("scripts/analysis/utils/filters.R"))
source(here("scripts/analysis/utils/stats.R"))
source(here("scripts/wrangle/tents.R"))
```

notes: 
- `lmerTest` docs: https://www.jstatsoft.org/article/view/v082i13

```{r load data, message=FALSE}
source(here::here("scripts/tidy/R/tents.R"))

# formatting
# imported_tents <- 
#   dfs_tents$n_pertent %>%
#   mutate(across(starts_with("trt"), as.factor))

#rm(dfs_pertent_ss)

# subset (all lab bugs only)
dfs <- list(

  # incl virgins; can use for laying but not hatching
  ## BUT note that for final viz, only looking at mated tents
  all = dfs_tidy$tents %>%
    filter(mate.pop != "field"),
  
  # use for hatching
  mated = dfs_tidy$tents %>% #dfs$tents
    filter(mate.pop != "field",
           mate.type != "virgin"),
  
  matedexptl = dfs_tidy$tents %>% #dfs$exptal
    filter(mate.pop != "field",
           mate.type != "virgin",
           mate.trt != 260)
)

# calc per tent SS and add cols/format
dfs <- dfs %>%
  lapply(., \(x){
    x %>%
      CalcTentCounts() %>%
      mutate(trt.minT = case_when(mate.trt == 419 ~ 19,
                                  mate.trt == 433 ~ 33,
                                  TRUE ~ 26),
             mate.ishs = factor(mate.ishs, levels = c("none", "both", "f", "m")),
             across(c("year", starts_with(c("trt", "mate"))), ~ as.factor(.x))
             )
  })

# make easier to interpret...
dfs$matedexptl$mate.trt <- dfs$matedexptl$mate.trt %>%
      recode_factor(`419` = "19", `426` = "26", `433` = "33")
```


# ## TRY 4: mss EDAs again

cuz now i have a better sense of what im doing...
(top = most recent)

## post-futz

per "futzing", make simpler subsets lol w just predictors of interest (this feels more right...)

```{r}
# 2023
df <- dfs$mated %>%
  filter(year == 2023)

mods <- list(
  all_win = glmer(n.coll ~ trt.type + (1|cage), offset = log(n.f),
                  family = "poisson", data = df %>% filter(mate.type == "within")),
  expt_win = glmer(n.coll ~ trt.minT + (1|cage), offset = log(n.f),
                   family = "poisson", data = df %>% filter(mate.type == "within" & trt.type == "expt")),
  
  all = glmer(n.coll ~ mate.ishs + (1|cage), offset = log(n.f),
              family = "poisson", data = df), # feels kinda wrong lol
  expt_btwn = glmer(n.coll ~ trt.minT + (1|cage), offset = log(n.f),
                    family = "poisson", data = df %>% filter(mate.type == "between")),
  expt_all = glmer(n.coll ~ trt.minT * mate.ishs + (1|cage), offset = log(n.f),
                   family = "poisson", data = df %>% filter(trt.type == "expt"))
)

# trt effects
drop1(mods$all_win, test = "Chisq") # no diff btwn ctrl/hs LOLOL
drop1(mods$expt_win, test = "Chisq") # no diff btwn hs groups

# sex effects
drop1(mods$all, test = "Chisq") # no diff if m/f/both/none
drop1(mods$expt_btwn, test = "Chisq") # no eff of hs
#drop1(mods$expt_all, test = "Chisq") # interaction is ns
drop1(update(mods$expt_all, ~ . - trt.minT:mate.ishs), test = "Chisq") # no effect of minT or hs!

## copying this over for 2025 into main res...
```

## futzing

```{r trying subsets}
# try to just subset the mating groups?
# (for performance, mostly lokoing at 'Score')
df <- dfs$mated %>%
  filter(year == 2023)

mods<- list(  
  poi_all = glmer(n.coll ~ mate.trt*mate.ishs + (1|cage), offset = log(n.f),
                  family = "poisson", data = df),
  poi_alla = glmer(n.coll ~ mate.trt + mate.ishs + (1|cage), offset = log(n.f),
                   family = "poisson", data = df),
  poi_all2 = glmer(n.coll ~ trt.type*mate.istrt + (1|cage), offset = log(n.f), # **
                  family = "poisson", data = df), 
  poi_all2a = glmer(n.coll ~ trt.type + mate.istrt + (1|cage), offset = log(n.f), # **
                    family = "poisson", data = df), 
  poi_all3 = glmer(n.coll ~ trt.ishs.f + trt.ishs.m + trt.minT + (1|cage), offset = log(n.f),
                  family = "poisson", data = df), 
  poi_all3a = glmer(n.coll ~ trt.f + trt.m + trt.type + (1|cage), offset = log(n.f),
                    family = "poisson", data = df)
)
check_model(mods$poi_all, panel = FALSE) %>% plot() # gr8 resids actually
check_model(mods$poi_alla, panel = FALSE) %>% plot() # gr8 resids actually
check_model(mods$poi_all2, panel = FALSE) %>% plot() # not the worst?
compare_performance(mods$poi_all,  mods$poi_alla, mods$poi_all2, mods$poi_all2a, mods$poi_all3, mods$poi_all3a, rank = TRUE) 
# 0.648, 0.588, 0.558, 0.648, 0.585, 0.617


mods<- list( 
  poi_expt = glmer(n.coll ~ mate.trt*mate.ishs + (1|cage), offset = log(n.f),
                   family = "poisson", data = df %>% filter(trt.type == "expt")),
  poi_expt2 = glmer(n.coll ~ trt.minT*mate.istrt + (1|cage), offset = log(n.f),
                  family = "poisson", data = df%>% filter(trt.type == "expt"))
)

check_model(mods$poi_expt, panel = FALSE) %>% plot() # not the worst, actually...
compare_performance(mods$poi_expt, mods$poi_expt2) # 0.687 (same models lol)


## win
mods <- list(
  poi_win = glmer(n.coll ~ mate.trt + (1|cage), offset = log(n.f),
                  family = "poisson", data = filter(df, mate.type == "within")),
  poi_win2 = glmer(sqrt(n.coll) ~ mate.trt + (1|cage), offset = log(n.f),
                   family = "poisson", data = filter(df, mate.type == "within"))
)

check_model(mods$poi_win, panel = FALSE) %>% plot() # bad var
check_model(mods$poi_win2, panel = FALSE) %>% plot() # meh var, doesnt do much?

compare_performance(mods$poi_win) # 0.565

## btwn
mods<- list(   
  poi_btwn = glmer(n.coll ~ mate.trt*mate.ishs + (1|cage), offset = log(n.f),
                   family = "poisson", data = df %>% filter(mate.type == "between")),
  poi_btwn2 = glmer(n.coll ~ mate.ishs + (1|cage), offset = log(n.f), # **
                   family = "poisson", data = df %>% filter(mate.type == "between"))
)

check_model(mods$poi_btwn, panel = FALSE) %>% plot() # also not that bad?

compare_performance(mods$poi_btwn, mods$poi_btwn2) # 0.796, 0.553)


## TLDR: think we need to subset lowkey LOL..(i.e. for w/in and btwn). also just compare to control simply
```


```{r eval = FALSE}
# lowkey hot garbage bc all the groups are combined?
mods <- list(
    poi = glmer(round(n.coll.perf) ~ 
                  mate.trt*mate.ishs + n.f + (1|cage),
              family = "poisson", 
              data = dfs$mated %>% filter(year == 2023)),
    poi_sq = glmer(round(sqrt(n.coll.perf)) ~ 
                  mate.trt*mate.ishs + n.f + (1|cage),
              family = "poisson", 
              data = dfs$mated %>% filter(year == 2023))
)

# mods <- list_modify(
#   mods,
    # poi_log = glmer(log(n.coll.perf+1) ~ 
    #               mate.trt*mate.ishs + n.f + (1|cage),
    #           family = "poisson", 
    #           data = dfs$mated %>% filter(year == 2023)),
#       nb = glmer.nb(round(n.coll.perf) ~ 
#                   mate.trt*mate.ishs + n.f + (1|cage),
#               data = dfs$mated %>% filter(year == 2023)),
#     nb_sq = glmer.nb(round(sqrt(n.coll.perf)) ~ 
#                   mate.trt*mate.ishs + n.f + (1|cage),
#               data = dfs$mated %>% filter(year == 2023))
# )

mod <- glmer(round(n.coll.perf) ~ 
                  mate.trt*mate.ishs + n.f + (1|cage),
              family = "poisson", 
              data = dfs$mated %>% filter(year == 2023))

# check
check_model(mods$poi, panel = FALSE) %>% plot()
check_model(mods$poi_sq, panel = FALSE) %>% plot()
#check_model(mods$poi_log, panel = FALSE) %>% plot()
check_model(mods$nb, panel = FALSE) %>% plot()
check_model(mods$nb_sq, panel = FALSE) %>% plot()

compare_performance(mods$poi, mods$poi_sq, mods$nb, mods$nb_sq) # log is busted

summary(mods$poi)


## OLD
# compare
anova(mods$poi, mods$nb, test = "Chisq")
## not diff, and AICpoi > AICnb, so stick with poisson?

# diagnose
mods$poi %>% DiagnoseModel2()


```


# ## TRY 3: EDAs for mss

couple of different explorations per feedback from july 2026 mtgs

2026-07-31 post-mortem: allowed to bench this now :)

## trying simpler models

2026-07-29 bc of all the imbalances in the numbers, just keep the predictors simple

```{r warning = FALSE}
# all data
df <- dfs$all %>%
  filter(year != 2024,
         mate.pop != "col")

mods <- list(
  m1 = glm(n.coll.perf ~ mate.type + n.f + year,
           family = "poisson", data = df),
  m2 = glm(n.coll.perf ~ trt.type + n.f + year,
           family = "poisson", data = df),
  m3 = glm(n.coll.perf ~ trt.type*mate.type + n.f + year,
           family = "poisson", data = df),
  m4 = glm(n.coll.perf ~ mate.trt*mate.istrt + n.f + year,
           family = "poisson", data = df)
)

anova(mods$m1, test = "Chisq") 
anova(mods$m2, test = "Chisq")
# differences for both mate/trt type... (so btwn ctrls/within are diff)

Anova(mods$m3, type = 2) # var order matters
# (everything matters...)

anova(mods$m4, test = "Chisq")
Anova(mods$m4, type = 2, 
      test.statistic = "LR",
      error.estimate = "deviance") 
  ## TODO idk what im doin here. does the LR Chisq col = the deviance col?
# everything matters...
```

## trying mixed models

trying cage as a random effect. (focusing on eggs collected first, then eggs hatched?)
eggs collected ~ # females + hs sex * hs trt + 1|enclosure

using 2023 eggs collected only (bc doing stats separately for each yr in mss), seeing 1) what response variable and 2) what distribution to use

```{r}
# df prep
# per mss, drop col stuff and 2024
y23 <- dfs$mated %>%
  filter(mate.pop != "col",
         year != 2024) %>%
  select(-c(mate.col, mate.pop, trt.pop)) %>%
  filter(year == 2023)

```

```{r}
# raw vs rounded poisson?
mods <- list(
  poi0 = glm(n.coll.perf ~ n.f + mate.trt*mate.ishs,
             data = y23, family = "poisson"),
  poi = glm(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs,
            data = y23, family = "poisson"),
  poioff = glm(n.coll ~ n.f + mate.trt*mate.ishs,
               data = y23, family = "poisson",
               offset = n.f)
)

summary(mods$poi0)
summary(mods$poi)

# diagnose them
anova(mods$poi0, test = "Chisq")
sum(resid(mods$poi0,type="pearson")^2)/df.residual(mods$poi0) # look at variance #69.08362
mods["poi0"] %>% DiagnoseModel2()

anova(mods$poi, test = "Chisq")
sum(resid(mods$poi,type="pearson")^2)/df.residual(mods$poi) #69.29671
## rounded looks pretty similar to the raw so just use the rounded

anova(mods$poioff, test = "Chisq")
sum(resid(mods$poioff,type="pearson")^2)/df.residual(mods$poioff) # 226.4761

AIC(mods$poi0, mods$poi, mods$poioff) # offset is not bueno
```

```{r}
# add the mixed eff model (and also cage as a fixed eff)
mods <- list_modify(
  mods,
  poic = glm(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + cage,
             data = y23, family = "poisson"),
  poim = glmer(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + (1|cage),
               data = y23, family = "poisson")
)

## compare them
AIC(mods$poi, mods$poic, mods$poim)
summary(mods$poic)
summary(mods$poim) # see fixed eff w/o cage

anova(mods$poim, mods$poic, mods$poi, test = "Chisq")
anova(moids$poi, mods$poim)

anova(mods$poim)
drop1(mods$poic, test = "Chisq")
drop1(mods$poim, test = "Chisq") # LRT on the fixed effs, per https://www.ssc.wisc.edu/sscc/pubs/MM/MM_TestEffects.html
anova(mods$poim, mods$poi, test = "Chisq") #idk wtf this means or if these mods are comparable TBH, since $poi isnt rly "nested" in $poim
mods[c("poi", "poic", "poim")] %>% DiagnoseModel2() #tbh the diagnosis is bad LOL -- plot vs resids is wonky

## so probably try fitting to an nb to see if the diagnostics look better?
```

```{r}
mods <- list_modify(
  mods,
  nb = glm.nb(n.coll.perf ~ n.f + mate.trt*mate.ishs,
              data = y23),
  nbc = glm.nb(n.coll.perf ~ n.f + mate.trt*mate.ishs + cage,
               data = y23), # will throw lots of warnings
  nbm = glmer.nb(n.coll.perf ~ n.f + mate.trt*mate.ishs + (1|cage),
                 data = y23)
)

AIC(mods$poi, mods$nb, mods$nbc, mods$nbm) # probably use nb model
mods[c("nb", "nbc", "nbm")] %>% DiagnoseModel2() #nb and nbm are better...(nbm is slightly better...?) nbc is too singular lol

## generally, still issues w/ singular fits
  # so maybe need to try the ctrl vs hot 2 grp thing joel suggested...
```

```{r}
mods <- list_modify(
  mods,
  poi_cve = glm(round(n.coll.perf) ~ n.f + trt.type,
                      data = y23, family = "poisson")
  )

anova(mods$poi_cve, test = "Chisq") # so there is a difference btwn ctrl/expt...
```

```{r}
mods <- list_modify(
  mods,
  poi_expt = glm(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs,
                 data = filter(y23, trt.type == "expt"),
                 family = "poisson"),
  poic_expt = glm(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + cage,
                  data = filter(y23, trt.type == "expt"),
                  family = "poisson"), # bad bc singular fits
  poim_expt = glmer(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + (1|cage),
                    data = filter(y23, trt.type == "expt"),
                    family = "poisson")
)

mods[c("poi_expt", "poic_expt", "poim_expt")] %>% DiagnoseModel2()
AIC(mods$poi_expt, mods$poim_expt)
  # reg poi is better diagnostics tho worse AIC
anova(mods$poim_expt, mods$poi_expt) # mixed mod is also better from anova..

anova(mods$poi_expt, test = "Chisq")
summary(mods$poim_expt) 
  # tho cant rly estimate variance when N<5 (see stats lab 9)... (i.e what we have here for tents lol)
  # https://stats.stackexchange.com/questions/601851/variance-estimation-for-small-sample-size
```


## EDA: effect of col

does effect of col matter?

2026-07-03: now ignoring this.. since per the viz it gets too piecemealy anyway

```{r}
# TODO wip
#' sth like:
#' 
#' eggs laid ~ trt.f*is.col.f + trt.m*is.col.m maybe?
```


# ## TRY 2: exptal only

per mtg feedback, focus on expt bugs only -- trying out transforming response variable (for eggs laid)

## try 2: (more) eggs laid transforms

TODO 
- try offsets for egg poissons?

- 2026-05-31:
 - trying out better model diagnosing first lol (for ntw mss final stats)
 - testing response variable transform + model type

### model type + Z
 
```{r warning = FALSE}
# look at overall distribution
hist(dfs$matedexptl$n.coll.perf, breaks = 25)
## right-skewed

# try basic poisson glm and check assumptions
mods <- list(
  poi = glm(n.coll.perf ~ n.f + mate.trt*mate.ishs + year,
            data = dfs$matedexptl, family = "poisson"),
  poi_sqrt = glm(sqrt.coll.perf ~ n.f + mate.trt*mate.ishs + year,
            data = dfs$matedexptl, family = "poisson"),
  poi_log = glm(log(n.coll.perf + 1) ~ n.f + mate.trt*mate.ishs + year,
            data = dfs$matedexptl, family = "poisson"),
  poi_round = glm(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + year,
            data = dfs$matedexptl, family = "poisson"),
  poi_rsqrt = glm(round(sqrt.coll.perf) ~ n.f + mate.trt*mate.ishs + year,
            data = dfs$matedexptl, family = "poisson")
)

# check residuals
mods %>% DiagnoseModel()
## sqrt and log+1 fix the resids 
## log is better than sqrt, tho some more outlier resids


# check for zero inflation (try the log+1)
## see: https://r-statistics.co/Zero-Inflated-Models-in-R.html
#obs_zeros = sum(dfs$matedexptl$n.coll.perf == 0)

# check overdispersion
## TODO: try performance::check_overdispersion? https://rdrr.io/cran/performance/f/README.md 

## refit to glm neg bin to use pscl::odTest
mods <- list_modify(
  mods,
  nb = glm.nb(n.coll.perf ~ n.f + mate.trt*mate.ishs + year,
              data = dfs$matedexptl),
  nb_sqrt = glm.nb(sqrt.coll.perf ~ n.f + mate.trt*mate.ishs + year,
                   data = dfs$matedexptl),
  # nb_rsqrt = glm.nb(round(sqrt.coll.perf) ~ n.f + mate.trt*mate.ishs + year,
  #                   data = dfs$matedexptl),
  nb_log = glm.nb(log(n.coll.perf + 1) ~ n.f + mate.trt*mate.ishs + year,
                  data = dfs$matedexptl)
)

mods$nb %>% DiagnoseModel()
mods$nb_sqrt %>% DiagnoseModel()
# mods$nb_rsqrt %>% DiagnoseModel()
mods$nb_log %>% DiagnoseModel()
## nb, sqrt, rsqrt diagnostic plots look about the same, so just use mods$nb
## log has best resids of the three

pscl::odTest(mods$nb)
## so use a neg bin dist...

# #################
# ### JKKKK ITS HATCH STUFF SHOULD BE ZERO INFLATED LOL not the eggs collected omg)
# 
# 
# # fit a zero-inflation model and compare to glmnb with AIC
# ## (dont need to transform response var for this...)
# # TODO: try transforming response var and other dists
# mods <- list_modify(
#   mods,
#   zip_nb = zeroinfl(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + year | 1,
#                     data = dfs$matedexptl, dist = "negbin"),
#   zip_poi = zeroinfl(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + year | 1,
#                     data = dfs$matedexptl, dist = "poisson"),
#   zip_nb2 = zeroinfl(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + year,
#                     data = dfs$matedexptl, dist = "negbin"),
# )
# 
# mods$zip_nb %>% DiagnoseModel2() 
# ## beautiful resids and QQ!!
# mods$zip_poi %>% DiagnoseModel2() 
# ## worse resids

# try some other dists before testing AICs
mods <- list_modify(
  mods,
  gau = glm(n.coll.perf ~ n.f + mate.trt*mate.ishs + year,
            data = dfs$matedexptl, family = "gaussian"),
  # Gamma is bad bc there are 0s
  # gam = glm(round(n.coll.perf) ~ n.f + mate.trt*mate.ishs + year,
  #           data = dfs$matedexptl, family = "Gamma"),
  qpoi = glm(n.coll.perf ~ n.f + mate.trt*mate.ishs + year,
            data = dfs$matedexptl, family = "quasipoisson")
)

# compare AICs
AIC(mods$gau, mods$qpoi, #mods$gam, # cant get LL for qpoi
    mods$poi_sqrt, mods$poi_log, # dont work bc non-integer Z
    mods$poi_round, mods$poi_rsqrt,
    mods$nb, mods$nb_sqrt, mods$nb_log  #mods$nb_rsqrt, # very similar to nb_sqrt
    #mods$zip_nb, mods$zip_poi # WHOOPS!
    )
## so lowkey just the negbin with log transform?

mods$nb_log %>% GetModelResults()
## nothin matters!!!!
```

```{r}
# eggs hatch (do the zero inflated here...)

```

### sep yrs?
```{r warning = FALSE}
# try doing stats per yr using the negbin(log+1)

mods <- list(
  all = glm.nb(log(n.coll.perf + 1) ~ n.f + mate.trt*mate.ishs + year,
                  data = dfs$matedexptl),
  all_noyr = glm.nb(log(n.coll.perf + 1) ~ n.f + mate.trt*mate.ishs,
                  data = dfs$matedexptl),
  y23 = glm.nb(log(n.coll.perf + 1) ~ n.f + mate.trt*mate.ishs,
                  data = dfs$matedexptl %>% filter(year == 2023)),
  y25 = glm.nb(log(n.coll.perf + 1) ~ n.f + mate.trt*mate.ishs,
                  data = dfs$matedexptl %>% filter(year == 2025))
)

# diagnose
mods %>% DiagnoseModel()
## all comparable to e/o. y25 on its own has the smallest resids

AIC(mods$all, mods$all_noyr, mods$y23, mods$y25) 
## model w/ all data is better if drop yr...
## smaller AICs in models w data subsets

mods %>% GetModelResults()
## nothin matters!!!
```


## try 1: eggs laid transforms

2026-05-31: archived trying out modeling eggs laid, focusing on just the expt bugs

### diagnostics + transforms

- tldr
  - offset... sometimes helps
  - need `n.f` as a predictor still.
  - rounding the rate is really nice!
  - so maybe use Z = round(eggs/F) for the models.

```{r warning=FALSE}
# Z: raw n.coll
# AICs: raw + offset = 4090.56; offset2 = 5013.118
mods <- list(
  raw = glm(n.coll ~ mate.trt*mate.ishs + n.f + year,
            data = dfs$matedexptl, family = "poisson"),
  # better QQ
  offset = glm(n.coll ~ mate.trt*mate.ishs + n.f + year, 
                   offset = n.f,
                   data = dfs$matedexptl, family = "poisson"),
  # bad QQ
  offset2 = glm(n.coll ~ mate.trt*mate.ishs + year, 
                    offset = n.f,
                    data = dfs$matedexptl, family = "poisson")
)

# Z: n.coll/f; sqrt(n.coll/f)
# AICs: none cuz poisonning non-integers
mods <- list(
  rawperf = glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
            data = dfs$matedexptl, family = "poisson"),
  # QQ looks like above
  offset =  glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
                        offset = n.f,
                        data = dfs$matedexptl, family = "poisson"),
  
  # (try the sqrt)
  # bad QQ
  sqrt_perf = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year,
             data = dfs$matedexptl, family = "poisson"),
  
  # QQ not that diff from 'offset'
  ## meh. this approach also doesnt feel right (using qb to deal with the non-integers)
  perf_qp = glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
                   data = dfs$matedexptl, family = "quasipoisson")
)

# Z: round(n.coll/f) (since dist is right-skewed with mean > 3)
# AICs: roundperf/offset = 1430.344, offset2 = 2531.161
mods <- list(
  roundperf = glm(round(n.coll.perf) ~ mate.trt*mate.ishs + n.f + year, # use this?
                  data = dfs$matedexptl, family = "poisson"),
  # worse QQ (slightly)
  offset = glm(round(n.coll.perf) ~ mate.trt*mate.ishs + n.f + year,
                   offset = n.f, 
                   data = dfs$matedexptl, family = "poisson"),
  # uhh worse resids vs fitted? (slopes down)
  offset2 = glm(round(n.coll.perf) ~ mate.trt*mate.ishs + year,
                   offset = n.f,
                   data = dfs$matedexptl, family = "poisson")
)

mods %>% DiagnoseModel()
mods %>% GetModelResults()
mods %>% lapply(., AIC)
```

### testing different pred vars
- using the mod: `glm(round(n.coll.perf) ~ mate.trt*mate.ishs + n.f + year, family = "poisson")` as the base

```{r eval = FALSE}
mods <- list(
  roundperf = glm(round(n.coll.perf) ~ mate.trt*mate.ishs + n.f + year, 
                  data = dfs$matedexptl, family = "poisson")
)

# NOTE: left incomplete... see try 2 instead
```


# ## TRY 1: all bugs (ref only)

- 2026-02-10 after wkly mtg, just for ref only. but TLDR:
  - `n.f` and `year` should be inluded in the eggs laid mods
    - 2026-04-20: also i think `year` was being treated as a number lol (or a char? at least not a factor...)
  - Z = `n.coll.perf` is actually prob fine as the response var? (i.e. dont need to transform)
    - BUT this gets weird w poisson bc of the non-integers (so still need to address this)
  - need to run on a subset of the data without the ctrls (or else just as ctrls vs hot bc 26-26 mating group is confounded w/ `mate.ishs` and `mate.trt`)
  - do the hatching as `c(# hatch, # coll)` (not as success/failures)

## eggs laid

### diagnosing Z

trying out different response variables for the egg numbers.

- `mods_laid$sqrt` is the better transform (smaller resids, better QQ but still a little off..)
- `mods_laid$log` has a very ugly QQ plot

```{r eval=FALSE, warning=FALSE}
# trying out difference response variables
#(<mod>2 = code the trt as a more straightfwd interaction)
mods <- list(
  raw = glm(n.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, 
            data = dfs$mated, family = "poisson"),
  raw2 = glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year, 
            data = dfs$mated, family = "poisson"),
  sqrt = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, 
             data = dfs$mated, family = "poisson"),
  sqrt2 = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year, 
             data = dfs$mated, family = "poisson"),
  log = glm(log(n.coll.perf + 1) ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, 
            data = dfs$mated, family = "poisson"),
  log2 = glm(log(n.coll.perf + 1) ~ mate.trt*mate.ishs + n.f + year, 
            data = dfs$mated, family = "poisson"),
  sqrt_q = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, 
             data = dfs$mated, family = "quasipoisson")
)

mods %>% DiagnoseModel() # use sqrt -- more balanced
```

```{r eval=FALSE, warning=FALSE}
# trying out different groupings
mods <- list(
  raw = glm(n.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, data = dfs$mated, family = "poisson"),
  sqrt = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, data = dfs$mated, family = "poisson"),
  log = glm(log(n.coll.perf + 1) ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.f + year, data = dfs$mated, family = "poisson")
)

mods %>% DiagnoseModel() # use sqrt -- more balanced
```

### trying pred vars

figuring out which vars to include in the model and what model to use
- Z: eggs laid/collected
  - testing pred vars
    - [x] number f: include
    - [x] year: include (as fixed bc only 2 yrs)
    - [x] cage (random/fixed?): dont include bc breaks
    - [ ] fridge
    - [ ] colony
  - [x] testing interactions:
    - code as `trt.f*trt.m + ishs.f*ishs.m`?
    
#### trts
- how to code the trts?

```{r eval=FALSE, warning=FALSE}
# lets give up on tents and look at the interactions instead
mods <- list(
  # which vars are needed?
  add = glm(sqrt.coll.perf ~ trt.f + trt.ishs.f + trt.m + trt.ishs.m,
            data = dfs$mated, family = "poisson"),
  add_nf = glm(sqrt.coll.perf ~ trt.f + trt.ishs.f + trt.m + trt.ishs.m + n.f,
            data = dfs$mated, family = "poisson"),
  add_yr = glm(sqrt.coll.perf ~ trt.f + trt.ishs.f + trt.m + trt.ishs.m + n.f + year,
              data = dfs$mated, family = "poisson"),
  
  # testing ints
  ## incl year and n.f
  int = glm(sqrt.coll.perf ~ trt.f:trt.ishs.f + trt.m:trt.ishs.m + n.f + year,
            data = dfs$mated, family = "poisson"),
  # (this is the "add" model)
  # int2 = glm(sqrt.coll.perf ~ trt.f*trt.ishs.f + trt.m*trt.ishs.m + n.f + year,
  #                data = dfs$mated, family = "poisson"),
  int_is = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.m:trt.ishs.f + n.f + year,
            data = dfs$mated, family = "poisson"),
  int_trt = glm(sqrt.coll.perf ~ trt.f:trt.m + trt.ishs.m:trt.ishs.f + n.f + year,
            data = dfs$mated, family = "poisson"),
  # (this is same as int_trt?)
  int_trt2 = glm(sqrt.coll.perf ~ trt.f*trt.m + trt.ishs.m*trt.ishs.f + n.f + year,
                 data = dfs$mated, family = "poisson"),
  int3 = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year,
             data = dfs$mated, family = "poisson")
)

# testing additional vars
anova(mods$add, mods$add_nf, test = "Chisq") # including n.f is better
anova(mods$add_nf, mods$add_yr, test = "Chisq") # including year is better
anova(mods$add, mods$add_yr, test = "Chisq")
drop1(mods$add_yr)
## so include yr and n.f

# testing add vs int?
anova(mods$add_yr, mods$int, test = "Chisq") # they are the same lol?

# testing coding interactions (trts + hs + sex separated)
anova(mods$int, mods$int_is, test = "Chisq") # int is better
anova(mods$int, mods$int_trt, test = "Chisq") # int_trt is marginally better
anova(mods$int, mods$int_trt2, test = "Chisq") # int_trt2 is marginally better

anova(mods$int_trt, mods$int_trt2, test = "Chisq") # these are the same? (so just use int_trt for the below)
anova(mods$int_is, mods$int_trt, test = "Chisq") # int_trt marginally better
anova(mods$add_yr, mods$int_trt, test = "Chisq") # int_trt marginally better
anova(mods$add_yr, mods$int3, test = "Chisq") # int3 marginally better...

# testing coding interactions (trts + hs separated)
anova(mods$int_trt, mods$int3, test = "Chisq") # these are the same?
AIC(mods$int_trt)
AIC(mods$int3)
drop1(mods$int_trt)
drop1(mods$int3)

# looking at the final model bc i think this is the better one?
mods$int3 %>% GetModelResults()

```

#### cage    
- works when fixed (and sig) but not rly what we're trying to do here...

```{r try 2, eval=FALSE, warning=FALSE}
mods <- list(
  nocage = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year,
               data = dfs$mated, family = "poisson"),
  cage_fixed = glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year + cage, # works but not what we want..
                   data = dfs$mated, family = "poisson")
  # breaks: bc of https://forum.posit.co/t/error-in-eval-f-x-downdated-vtv-is-not-positive-definite/73846/2?
  # cage_rand = glmer(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year + (1|cage),
  #                  data = dfs$mated, family = "poisson")
)

mods %>% GetModelResults()

mods %>% DiagnoseModel()
```

- `cage` breaks things. not rly sure why though... (colinearity?)

```{r try 1 cages break, eval=FALSE, warning=FALSE, message=FALSE}
# drop tents with only 1 m/f trt combo (colinear?) for diagnosing
dfs <- list_modify(
  dfs,
  mod = dfs$mated %>%
  group_by(trt.f, trt.m) %>%
  mutate(n = n(),
         cage = case_when(n == 1 ~ NA_character_,
                          TRUE ~ cage))
  )

mods <- list(
  # diagnosing mods
  # full = glm(sqrt.coll.perf ~ mate.trt*trt.ishs.f*trt.ishs.m + n.f + cage + year,
  #            data = dfs$mated, family = "poisson"),
  # cage_fixed = glm(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.m + trt.ishs.f + n.f + cage,
  #              data = dfs$mod, na.action = na.omit,
  #              family = "poisson"),
  # cage_rand = glmer(sqrt.coll.perf ~ trt.f + trt.m + trt.ishs.m + trt.ishs.f + n.f + (1|cage),
  #            data = dfs$mod, na.action = na.omit,
  #            family = "poisson"),
  test = glm(sqrt.coll.perf ~ trt.f + trt.m,
             data = dfs$mated, family = "poisson")
  
)
  
mods %>% GetModelResults()
mods %>% DiagnoseModel() # breaks (NA/missing data) 

resid(mods$test) # try -- when `cage` is included, resids are 0

```

### trying stuff w joel
- trying stuff with the mods:
  - use just a subset of the hot bugs only
  - try looking at the trt interaction + dropping different terms
  - trying trt as numeric vs factor
  - try other distributions

```{r trying stuff w joel 260210, eval=FALSE}
# how are the stats when the ctrl group is dropped?
df <- dfs$mated %>% filter(mate.trt != 260)

# if Z = n.coll these are amazing btw
mods <- list(
  # try dropping different terms
  m1 = glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
            data = df, family = "poisson"),
  m2 = glm(n.coll.perf ~ mate.trt + mate.ishs + n.f + year,
            data = df, family = "poisson"),
  m3 = glm(n.coll.perf ~ mate.ishs + n.f + year,
            data = df, family = "poisson"),
  # try trt as numeric?
  m4 = glm(n.coll.perf ~ as.numeric(mate.trt)*mate.ishs + n.f + year,
            data = df, family = "poisson"),
  # fyi: different dist from m1-4 so cant anova it with the rest
  m5 = glm(n.coll.perf ~ mate.ishs + n.f + year,
            data = dfs$mated, family = "poisson")
)

mods %>% GetModelResults()

anova(mods$m1, mods$m2, mods$m3, mods$m4, test = "Chisq")
AIC(mods$m1, mods$m2, mods$m3, mods$m4)

anova(mods$m5, test = "Chisq")
summary(mods$m5)

# trying other transforms/dists
glm(n.coll.perf ~ mate.trt*mate.ishs + n.f + year,
            data = df, family = "gaussian") %>% DiagnoseModel()
glm(sqrt.coll.perf ~ mate.trt*mate.ishs + n.f + year,
            data = df, family = "gaussian") %>% DiagnoseModel() # kinda bad
glm(n.coll.perf + 1 ~ mate.trt*mate.ishs + n.f + year,
            data = df, family = "Gamma") %>% DiagnoseModel()
```


## eggs hatched

### diagnosing Z

- dont need `n.f` in the model...? tho maybe need `n.coll.perf`?
- they all REALLY suck... need a different model? drop that one ctrl point?
  - try zero-inflated poissons!
  
#### glms (bad: bc dist)
```{r , eval=FALSE, warning=FALSE, message=FALSE}
# i also just think this is bad bc of the preds LOL...
# regular glms:
mods <- list(
   raw = glm(n.hatch.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.coll.perf + year, data = dfs$mated, family = "poisson"),
  # sqrt = glm(sqrt.hatch.perf ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.coll.perf + year, data = dfs$mated, family = "poisson"),
  # cbrt = glm((n.hatch.perf)^(1/3) ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.coll.perf + year,
  #            data = dfs$mated, family = "poisson"),
  log = glm(log(n.hatch.perf + 1) ~ trt.f + trt.m + trt.ishs.f + trt.ishs.m + n.coll.perf + year, data = dfs$mated, family = "poisson")
)

mods %>% DiagnoseModel() # they all really suck (cube rt is sorta okay)
qqnorm(mods$log$residuals); qqline(mods$log$residuals)
qqnorm(resid(mods$log)); qqline(resid(mods$log))
```

```{r eval=FALSE, warning=FALSE}
# more glms with different response var and diff preds
# (try on the whole number instead of hatch per f bc that doesn't rlly make sense... also try w the trt interaction)

# glm breaks vars are when factors
df <- dfs$mated %>%
  mutate(across(c("year", starts_with(c("mate", "trt"))), as.character))

mods <- list(
  raw = glm(n.hatch ~ mate.trt*mate.ishs + year,
            data = df, family = "poisson"),
  log = glm(log(n.hatch + 1) ~ mate.trt*mate.ishs + year, 
            data = df, family = "poisson")
  # raw_nb = MASS::glm.nb(n.hatch ~ mate.trt*mate.ishs + year,
  #              data = dfs$mated) # errors...
)

mods %>% DiagnoseModel()
```

#### zips (try 1)
- need to code the trts differently.. otherwise warnings

```{r eval=FALSE}
# zip/zinb
mods <- list(
  raw_p = zeroinfl(n.hatch ~ trt.f|1, data = dfs$mated),
  raw_nb = zeroinfl(n.hatch ~ trt.f|1, data = dfs$mated, dist = "negbin")
  # raw_p2 = zeroinfl(n.hatch ~ mate.trt*mate.ishs + year, data = dfs$mated),
  # raw_nb2 = zeroinfl(n.hatch ~ mate.trt*mate.ishs + year, data = dfs$mated, dist = "negbin")
  
  # break: probably bc of how the trt info is handled
  # raw_p = zeroinfl(n.hatch ~ trt.f + trt.m + as.factor(trt.ishs.f) + as.factor(trt.ishs.m) + n.coll + year, 
  #                data = dfs$mated),
  # raw_nb = zeroinfl(n.hatch ~ trt.f + trt.m + as.factor(trt.ishs.f) + as.factor(trt.ishs.m) + n.coll + year, 
  #                data = dfs$mated, dist = "negbin")

)


mods %>% DiagnoseModel2()
#qqnorm(mods$raw_p$residuals); qqline(mods$raw_p$residuals)
#qqnorm(resid(mods$raw_p)); qqline(resid(mods$raw_p))
```

#### zips (try 2)
- zip model wont fit if the trts are too piecemealy, so try just `mate.trt * mate.ishs`
  - (try `year` and `n.f` later)

```{r eval=FALSE}
df <- dfs$mated %>%
  mutate(mate.ishs = case_when(mate.ishs == "none" ~ NA_character_,
                               TRUE ~ as.character(mate.ishs)))

mods <- list(
  raw_p = zeroinfl(n.hatch ~ mate.trt*mate.ishs|1, 
                   data = df, dist = "poisson"),
  raw_nb = zeroinfl(n.hatch ~ mate.trt*mate.ishs|1, 
                    data = df, dist = "negbin")
)

mods %>% DiagnoseModel2() # resids are a little tighter w the negbin. does it need to be transformed tho ??

```

### trying pred vars

- where Z: eggs hatched
  - (glms are sorta wonky bc the distribution is sooo bad to begin w lol... (try ZIP model instead))
  - nts: dont need number of F for this.. just # of eggs coll???

## binomial (hatch, laid)

- [wip] Z: binomial (ie # hatch, # coll)
  - wonkkkyyy...

### trying pred vars
- seems like `trt` of the m/f is more impt? (everything sig except `trt.ishs`...)
- but go look at data transforming before continuing more lol

```{r, eval=FALSE}
# for simplicity, focus on the trts
dfs <- list_modify(
  dfs,
  mods = dfs$mated %>%
    mutate(n.fail = n.coll - n.hatch)
)

mods <- list(
  # as (# successes, # failures)
  raw_sf = glm(cbind(n.hatch, n.fail) ~
                 #trt.f + trt.m + trt.ishs.m + trt.ishs.f + n.f + year,
                 #trt.f*trt.m + trt.ishs.f*trt.ishs.m + n.f + year,
                 mate.trt*mate.ishs + n.f + year,
               data = dfs$mods, family = "binomial"),
  
  
  # as a prop
  raw_hc = glm(cbind(n.hatch, n.coll) ~
                 #trt.f + trt.m + trt.ishs.m + trt.ishs.f + n.f + year,
                 #trt.f*trt.m + trt.ishs.f*trt.ishs.m + n.f + year,
               mate.trt*mate.ishs + n.f + year,
               data = dfs$mods, family = "binomial"),
  
  raw_hprop = glm(n.hatch/n.coll ~ mate.trt*mate.ishs + n.f + year,
                  #trt.f*trt.m + trt.ishs.f*trt.ishs.m + n.f + year,
               data = dfs$mods, family = "quasibinomial") # use qb for non-int binomials
)

#mods %>% DiagnoseModel() # breaks bc of cbind
mods %>% GetModelResults()
#anova(mods$raw_sf, mods$raw_hc, test = "Chisq") # breaks bc of cbind?
```
# ### misc

## sample sizes

```{r}
dfs_tidy$tents %>%
  CalcTentCounts() %>% 
  filter(!is.na(trt.f),
         !is.na(trt.m), # drops virgins
         mate.pop != "col" # drops col ctrl matings
         ) %>%
  # treating col as lab bugs ?
  mutate(mate.pop = case_when(mate.pop == "col" ~ "lab",
                              TRUE ~ mate.pop)) %>% #View()
  group_by(across(c("year", "mate.type", "mate.pop",
                    "trt.f", "trt.m"))) %>%
  summarise(total.enclosures = n(),
            total.females = sum(n.f),
            total.males = sum(n.m),
            total.eggs = sum(n.coll),
            total.hatch = sum(n.hatch)
            )
```

