---
title: "ntw dev data tidying"
date: "2025-11-26"
---

# # preamble

tidy data from all ntw-related expts for wrangling/analyses.

(data pre-cleaning is somewhat inconsistent across the years, so just standardise everything here)

final code from this `.Rmd` is purled to a `.R` script; rerun `purl` step as needed to update the wrangling for analyses. outputs data from all expts to keep it all standard, but load each data subset into its analysis script with the corresponding `load_*.R` script.


```{r set paths, eval=FALSE, include=FALSE, purl=FALSE}
# run this prior to other pathing stuff if things are breaking
here::i_am("scripts/tidy-dev.Rmd")
```

```{r for purling, eval=FALSE, include=FALSE, purl=FALSE}
# rerun this chunk if purled .R script needs to be updated.
# save and apply updates to current .Rmd before purling!

knitr::purl("tidy-dev.Rmd", here::here("scripts/R/tidy-dev.R"))
```

# load libs & pkgs

```{r about}
# cleaning/ntw-2025.R

# knitted tidying code for 2025 ntw dev figs/analyses.
# source() this in corresponding analysis scripts.

# this script compiles ntw data from 2023-25 into cohesive units for easier wrangling.
```

```{r load libs}
library(data.table) # for rbindlist
library(tidyverse)
library(here)
```

# # dev data (all yrs)

## loading
```{r load data}
dfs_imported <- list(
  y23 = read.csv(here("data/clean/2023/dev.csv"), header = TRUE) %>%
    # dealing w dup ids
    mutate(id = as.numeric(id), 
           id = case_when(id == 1379 & !is.na(jdate.pmd) ~ 1379.1,
                          #expt.group == "A" ~ paste0(id, "a"),
                          expt.group == "A" ~ id + 0.1,
                          TRUE ~ id),
           # handle this way earlier
           trt = case_when(trt == 337 ~ 426,
                           TRUE ~ as.numeric(trt))),
    #mutate(id = sprintf("%.1f", id)),
  y24 = read.csv(here("data/clean/2024/dev.csv"), header = TRUE),
  y25 = read.csv(here("data/clean/2025/dev.csv"), header = TRUE),
  a23 = read.csv(here("data/clean/2023/tent-matings.csv"), header = TRUE) %>%
    select(c("id", "trt", "sex", "pop", "id.tent",
             "jdate.ec", "jdate.paired", "jdate.died", "ctrl", "if.mated", "if.fridge")) %>%
        filter(!(id %in% c(282, 284))), # bad bugs
  a24 = read.csv(here("data/clean/2024/tent-matings.csv"), header = TRUE) %>%
    select(c("id", "trt", "pop", "sex", "cage", 
             "jdate.tent", "if.mated", "if.fridge")),
  a25 = read.csv(here("data/clean/2025/tent-matings.csv"), header = TRUE) %>%
    select(c("id", "trt", "sex", "pop", "cage",
             "jdate.eclosed", "jdate.added", "jdate.removed", 
             "if.fridge", "if.def", "if.culled", "if.mated")) %>%
    filter(is.na(if.culled)) %>% # dont rlly wanna deal w these 2 bugs
    select(-if.culled)
)
```

## standardising

### initial filter and column dropping:
- drop MQs, "ignore" bugs
- hatchlings only

```{r filtering and column cleanup}
dfs_filtered <- list(
  # larva
  ## preremove all MQs
  y23 = dfs_imported[["y23"]] %>%
    filter(is.na(species),
           !is.na(instar.enter), # drop non-expt bugs
           # instar.enter == "hatch" | 
           #   pop == "field" & instar.enter %in% c("1st", "2nd"),
           #pop %in% c("field", "lab"), # drop F1s
           #!(trt %in% c("ctrl", "267", "330")),
           !(reason.ignore %in% c("lost", "early field")) |
             is.na(reason.ignore),
           ) %>%
    rename(#id = ID,
           #jdate.LPI = jdate.LP,
           #trt = treatment,
           cohort = expt.group,
           jdate.long = jdate.surv) %>%
    select(-c(starts_with(c("h.", "if.")), ends_with("2nd"), 
              #"instar.enter",  
              "jdate.collected", "jdate.enter", 
              "jdate.stuck", "jdate.15", "jdate.died",
              #"parent.tent", "src",
              "trt.stage", "grp.trt")),

  y24 = dfs_imported[["y24"]] %>%
    filter(is.na(species),
           instar.enter == "hatch") %>%
    rename(jdate.long = jdate.surv) %>%
    select(-c(ends_with("2nd"), "jdate.enter", #"instar.enter", 
              "trt.type")),
  
  y25 = dfs_imported[["y25"]] %>%
    filter(is.na(species),
           !(notes.ignore == "empty") | is.na(notes.ignore)) %>%
    rename(pop = population,
           jdate.long = jdate.surv,
           cage = id.cage) %>%
    mutate(cage = NA) %>% # cage ids are different in adults
    select(-c("species", "jdate.enter", #"instar.enter", 
              ends_with("2nd"))),
  
  # adults
  a23 = dfs_imported[["a23"]] %>%
    #mutate(if.fridge = case_when(jdate.ec < jdate.paired ~ "x")) %>%
    # mutate(if.fridge = case_when(!is.na(jdate.ec) & 
    #                                (jdate.ec < jdate.paired) ~ "x")) %>%
    #filter(!is.na(jdate.died)) %>% 
    rename(jdate.eclose = jdate.ec,
           jdate.tent = jdate.paired,
           jdate.long = jdate.died,
           trt.ctrl = ctrl,
           cage = id.tent),
  
  a25 = dfs_imported[["a25"]] %>%
    rename(jdate.tent = jdate.added,
           jdate.eclose = jdate.eclosed,
           jdate.long = jdate.removed)
)

# fix ids of supplemental mating bugs
dfs_filtered[c("a23", "a25")] <- dfs_filtered[c("a23", "a25")] %>%
  lapply(., \(x){
    x %>%
      mutate(id = case_when(
        grepl("-m", id) ~ 3100 + as.numeric(str_sub(id, start = -2)),
        grepl("-f", id) ~ 3200 + as.numeric(str_sub(id, start = -2)),
        TRUE ~ as.numeric(id)),
      )
  })
```

### merge adult/larval dev data

do in two steps (bc the col ctrls break `coalesce()` bc "missingness is uneven"):
1. `left_join` larval dev onto existing adult dev
2. `merge` remaining col adults into the dev data
```{r merging dfs by yr}
# merge adult + larval data
dfs_stdised <- list(
  y23 = dfs_filtered[["y23"]] %>% # 716+
    left_join(filter(dfs_filtered[["a23"]], id %in% dfs_filtered$y23$id), 
              by = c("id", "trt", "sex", "pop")) %>% #View()
    mutate(#jdate.tent = coalesce(jdate.tent.x, jdate.tent.y),
           jdate.eclose = coalesce(jdate.eclose.x, jdate.eclose.y),
           jdate.long = coalesce(jdate.long.x, jdate.long.y),
           .keep = "unused") %>% 
    merge(., filter(dfs_filtered[["a23"]], !(id %in% dfs_filtered$y23$id)), 
          all = TRUE) %>%
    mutate(year = 2023,
           #if.fridge = case_when(jdate.eclose < jdate.tent ~ "x")
           ),
  
  y24 = dfs_filtered[["y24"]] %>%
    left_join(filter(dfs_imported[["a24"]], id %in% dfs_imported$y24$id), 
              by = c("id", "trt", "sex", "pop")) %>%
    mutate(year = 2024),
  
  y25 = dfs_filtered[["y25"]] %>%   # 752+
    left_join(filter(dfs_filtered[["a25"]], id %in% dfs_filtered$y25$id), 
              by = c("id", "trt", "sex", "pop"#, "cage"
                     )) %>%
    mutate(jdate.eclose = coalesce(jdate.eclose.x, jdate.eclose.y),
           jdate.long = coalesce(jdate.long.x, jdate.long.y),
           cage = coalesce(cage.x, cage.y),
           .keep = "unused") %>% 
    merge(., filter(dfs_filtered[["a25"]], !(id %in% dfs_filtered$y25$id)), 
          all = TRUE) %>%
    mutate(year = 2025)
  
) %>% lapply(., \(x){
  # stdise column types
  x %>%
    mutate(across(c("id", "trt", starts_with("jdate"), ends_with("T", ignore.case = FALSE)),
           as.numeric))
})
```

### standardising values + matching columns for merging

cols that mean different things for different yrs:
- `trt.ctrl`: ctrl bug type
  - 2023: if diurnal ctrl or not
  - 2025: if colony ctrl or not
- `is.ctrl`: if colony bug or not (for mating stuff)
  
individual standardising of cols:
- dev cols thru 8th
- `trt.<...>`: for filtering trt stuff
- `fate`/survival stuff: dev outcome relative to pup (pmd/culled/ec/pup/other)
- handling dates for culled/exiting

```{r standardising values by yr}
# individual fixes: recoding values + adding missing columns
# stdising date meanings:
  # pmd = died before pup/LPI -> finalise in next section
  # cull = non-pmd larval death
  # surv = death after eclosion

# 2023: redo trt values, survival coding/notes, sup coding; add trt group
dfs_stdised[["y23"]] <- dfs_stdised[["y23"]] %>% 
  mutate(mass.8th = NA, jdate.8th = NA, 
         # if.fridge = case_when(!is.na(jdate.tent) & jdate.tent - jdate.eclose > 0 ~ "x"), 
         ## calc not relevant, just need the col: 2023 fridge bugs alrdy excluded from longev calcs
         trt.flucT = case_when(trt == 260 ~ 0, # force diurnals
                               TRUE ~ trt.flucT),
         trt.group = case_when(pop == "lab" & diet == "TB" ~ 1000,
                               pop == "lab" & diet == "LD" ~ 2000,
                               pop == "field" ~ 100),
         fate.dev = case_when(!is.na(jdate.eclose | pop == "col") ~ "ec",
                              !is.na(jdate.pupa) | !is.na(jdate.LPI) ~ "pup",
                              reason.ignore %in% c("wet diet", "accidental", "hot larva", "cut") | 
                                final.fate %in% c("accidental", "culled") ~ "other",
                              !is.na(jdate.pmd) ~ "pmd"),
         # jdate.culled = case_when(final.fate %in% c("accidental", "culled") & 
         #                            fate.dev != "pup" ~ as.numeric(jdate.exit)),
         jdate.culled = case_when(fate.dev == "other" ~ as.numeric(jdate.exit),
                                  TRUE ~ NA_real_),
         jdate.pmd = case_when(!is.na(jdate.culled) ~ NA_real_,
                               TRUE ~ jdate.pmd), 
         is.ctrl = case_when(#trt.flucT == 2.5 ~ "diurnal",
                             pop == "col" ~ "col")
         ) %>% 
  select(-c("final.fate", "reason.ignore"))

# 2024: redo survival coding, trt group
dfs_stdised[["y24"]] <- dfs_stdised[["y24"]] %>% 
  mutate(jdate.LPI = NA, 
         parent.tent = NA, src = NA,
         jdate.exit = NA, is.ctrl = NA,
         trt.group = case_when(id < 1000 ~ 100,
                               id > 2000 ~ 2000,
                               TRUE ~ 1000),
         fate.dev = case_when(!is.na(jdate.eclose) ~ "ec",
                              !is.na(jdate.pupa) ~ "pup",
                              !is.na(jdate.culled) | surv.outcome == 2  ~ "other",
                              !is.na(jdate.pmd) ~ "pmd"),
         jdate.culled = NA,
         jdate.long = case_when(is.na(jdate.eclose) ~ NA_real_,
                                TRUE ~ as.numeric(jdate.long)),
         #if.mated = case_when(!is.na(jdate.long) ~ 0)
         ) %>%
  select(-"surv.outcome")

# 2025: redo survival coding/notes; add timepts, col ctrl info (per 2023)
dfs_stdised[["y25"]] <- dfs_stdised[["y25"]] %>% 
  mutate(parent.tent = NA, src = NA,
         fate.dev = case_when(!is.na(jdate.eclose) & !is.na(jdate.long) ~ "ec",
                              !is.na(jdate.pupa) | !is.na(jdate.LPI) ~ "pup",
                              !is.na(jdate.culled) | !is.na(notes.ignore)  ~ "other",
                              !is.na(jdate.pmd) ~ "pmd"),
         jdate.culled = case_when(fate.dev == "other" ~ as.numeric(jdate.pmd),
                                  !is.na(jdate.pupa) | !is.na(jdate.LPI) ~ NA_real_,
                                  TRUE ~ jdate.culled),
         jdate.pmd = case_when(!is.na(jdate.culled) ~ NA_real_,
                               TRUE ~ jdate.pmd), 
         is.ctrl = case_when(pop == "col" ~ "col")) %>% 
  filter(!is.na(status) | id > 3000) %>%
  select(-c(starts_with("notes"), "status"))
```

simultaneous standardising:

(i.e. cols existed already, just making them all match up here)

- `is.<...>`: if is, what it is
- `jdate.exit` (bc all the dates should be standard now lol.)

```{r standardisng values for all}
# combine dfs; remove some short-lived larvae; more standardising:
## fix: exit/pmd dates; LPI info; late eclosions
## redo: survival, pmd/pup/ec/sup coding (counts after reaching 5th)
## add parsing: non-ntw trts, expt vs control temp trts
df_wide <- 
  data.table::rbindlist(dfs_stdised, fill = TRUE) %>% 
    mutate(
      # survival/pup/sup/ec status fixes
      # is.pup = case_when(fate.dev %in% c("ec", "pup") ~ 1,
      #                            fate.dev == "pmd" ~ 0,
      #                            fate.dev == "other" ~ 2),
      jdate.pmd = case_when(!is.na(jdate.pupa) | !is.na(jdate.LPI) ~ NA_real_,
                            TRUE ~ jdate.pmd),
      is.pup = case_when(!is.na(jdate.pmd) ~ 0,
                         !is.na(jdate.pupa) ~ 1,
                         !is.na(jdate.LPI) ~ 2),
      is.LPI = case_when(!is.na(jdate.LPI) ~ 1),
      # is.ec = case_when(fate.dev == "ec" ~ 1,
      #                   !is.na(jdate.pupa) & is.na(jdate.LPI) & 
      #                     is.na(jdate.eclose) ~ 0),
      ## only count expt bugs for eclosions
      is.ec = case_when(!is.na(instar.enter) & !is.na(jdate.pupa) & is.na(jdate.eclose) ~ 0,
                        !is.na(instar.enter) & !is.na(jdate.eclose) ~ 1
                        ),
      sup = case_when(!is.na(jdate.8th) ~ 8,
                      !is.na(jdate.7th) ~ 7,
                      !is.na(jdate.6th) ~ 6,
                      #!is.na(jdate.5th) ~ 5,
                      #FALSE ~ as.numeric(sup)
                      ),
      is.sup = case_when(!is.na(sup) ~ 1,
                         !is.na(jdate.wander) & is.na(sup) ~ 0),
      # LPI info fixing
      sex = case_when(!is.na(jdate.LPI) ~ NA_character_,
                      TRUE ~ sex),
      mass.pupa = case_when(!is.na(jdate.LPI) ~ NA_real_,
                            TRUE ~ mass.pupa),
      # late eclosion fixes (dates/masses are inaccurate)
      ## (based off the longest tt.eclose with longevity data/checking spreadsheet)
      jdate.eclose = case_when(jdate.eclose - jdate.pupa > 70 ~ NA_real_,
                               TRUE ~ jdate.eclose),
      mass.eclose = case_when(is.na(jdate.eclose) ~ NA_real_,
                              TRUE ~ mass.eclose), 
      # exit date fixes
      jdate.exit = case_when(!is.na(jdate.long) ~ as.numeric(jdate.long),
                             !is.na(jdate.eclose) ~ as.numeric(jdate.eclose),
                             !is.na(jdate.LPI) ~ as.numeric(jdate.LPI),
                             # !is.na(jdate.pupa) & is.na(jdate.eclose) ~
                             #   as.numeric(jdate.pupa + 36), # 15d + 3 wks
                             !is.na(jdate.culled) ~ as.numeric(jdate.culled),
                             !is.na(jdate.pmd) ~ as.numeric(jdate.pmd)
                             ),
      # more trt binaries
      # is.ntw = case_when(trt %in% c(419, 433) ~ 1,
      #                    trt %in% c(260, 426) ~ 2,
      #                    TRUE ~ 0),
      trt.type = case_when(trt.maxT == 40 ~ "expt",
                           TRUE ~ "ctrl")
      ) %>%
    # drop field bugs that die within 1 day
    filter(jdate.exit - jdate.hatch > 1 | is.na(jdate.hatch) | is.na(jdate.exit))

# add in colony bug trt info
df_wide[df_wide$pop == "col", c("trt.minT", "trt.maxT", "trt.meanT")] <- 26
df_wide[df_wide$pop == "col", "trt.flucT"] <- 0
```

# # prep for final export

## collate
```{r add calcs for pivot}
df_wide <- df_wide %>%
  select(-species) %>%
  mutate(tt.3rd = jdate.3rd - jdate.hatch,
           tt.4th = jdate.4th - jdate.hatch,
           tt.5th = jdate.5th - jdate.hatch,
           tt.6th = jdate.6th - jdate.hatch,
           tt.7th = jdate.7th - jdate.hatch,
           tt.8th = jdate.8th - jdate.hatch,
           tt.wander = jdate.wander - jdate.hatch,
           tt.pupa = case_when(!is.na(jdate.pupa) ~ jdate.pupa - jdate.hatch,
                               !is.na(jdate.LPI) ~ jdate.LPI - jdate.hatch),
           tt.eclose = jdate.eclose - jdate.pupa,
           #tt.long = case_when(is.na(if.fridge) ~ jdate.long - jdate.eclose),
           tt.long = jdate.long - jdate.eclose,
            ## for fridge bugs: jdate.long - jdate.tent for them? or jdate.exit?
            ## but some bugs were culled the day of, so jdate.eclose = jdate.exit?
           #tt.fridge = case_when(!is.na(if.fridge) ~ jdate.tent - jdate.eclose)
            ## TODO omit tt.fridge? not that useful. could calculate on the fly later
           #tt.exit = jdate.exit - jdate.hatch
            ## NOTE omit: just add this in later or else sorta interferes w PrepLarvalSS
    )

```


## pivot fn
```{r}
tolong <- function(x){
  x %>%
    pivot_longer(cols = starts_with(c("jdate", "mass", "tt")),
                 names_to = c(".value", "instar"),
                 values_drop_na = TRUE,
                 names_pattern = ("([a-z]*\\d*)\\.(\\d*[a-z]*)")) %>%
    
    # # drop empty dev times but keep masses
    # mutate(tt = case_when(is.na(tt) & !is.na(mass) ~ 999,
    #                       TRUE ~ as.numeric(tt))) %>%
    # drop_na(tt) %>%
    # mutate(tt = na_if(tt, 999)) %>%
  
  # drop longevity durations for short-lived eclosed bugs/culled on eclosion
  filter(!(instar == "long" & tt == 0))
}
  
```

# # cleanup
```{r remove intermediate objects}
rm(dfs_imported, dfs_filtered, dfs_stdised)
```

